### 2022-03-18 outputs/paper/paper.Rmd
...finally

---
title: "Numerous variables relating to 2016 election in 2021 US GSS"
subtitle: "Possible factors effecting 2016 election"
author: 
  - SangWoong Lee
  - Young Suk 
thanks: "Code and data are available at: https://github.com/SANGWOONG-LEE/2016electionanalysis.git".
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

# Introduction

Donald Trump won the 2016 United States presidential election. We are not US citizens. 
However, considering

# Data
1.Conveying a sense of the dataset of interest.
2.Dicussion of the survey's methodology and its key features, strengths, weaknesses.
For instance what is the population, frame and sample.
How is the sample recruited 
what sampling approach is taken
what are some of the trade offs of this 
how is non response handled.
(What is good and bad about it)

# Results
Convey findings.

# Discussion

Things that we must include

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Supplementary survey

Our supplementary survey is available here: LINK

## Preamble:

The purpose of this survey is ~~

By proceeding with this survey, you understand that we will user your responses to 
better understand () in the GSS. The survey is voluntary, and if you decide to 
participate you can skip questions and withdraw at any time.

## Questions
1. 
2. 
3.

## End Page

Thank you ~~

\newpage

# References

### 2022-03-18 outputs/paper/paper.Rmd
ggplot

@@
 (What is good and bad about it)
+
+```{r}
+library(ggplot2)
+reduced_data %>%
+  ggplot(aes(x = gender, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = degree, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = born_US, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = Family_income, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = nation_education, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = nation_race, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = nation_arms, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = nation_welfare, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = Assistance_to_the_poor, fill = pres16)) +
+  geom_bar()
+```
 

### 2022-03-18 outputs/paper/paper.Rmd
,,,

--- 2022-03-18 outputs/paper/paper.Rmd
@@
 (What is good and bad about it)
+
+```{r}
+library(ggplot2)
+reduced_data %>%
+  ggplot(aes(x = gender, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = degree, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = born_US, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = Family_income, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = nation_education, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = nation_race, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = nation_arms, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = nation_welfare, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = Assistance_to_the_poor, fill = pres16)) +
+  geom_bar()
+```
+```{r}
+reduced_data %>%
+  ggplot(aes(x = Assistance_to_other_countries, fill = pres16)) +
+  geom_bar()
+```
 

### 2022-03-18 outputs/paper/paper.Rmd
no change
ggplot

--- 2022-03-18 outputs/paper/paper.Rmd
@@
 (What is good and bad about it)
+
+```{r}
+library(ggplot2)
+reduced_data %>%
+  ggplot(aes(x = gender, fill = pres16)) +
+  geom_bar()
+
+reduced_data %>%
+  ggplot(aes(x = degree, fill = pres16)) +
+  geom_bar()
+
+reduced_data %>%
+  ggplot(aes(x = born_US, fill = pres16)) +
+  geom_bar()
+
+reduced_data %>%
+  ggplot(aes(x = Family_income, fill = pres16)) +
+  geom_bar()
+
+reduced_data %>%
+  ggplot(aes(x = nation_education, fill = pres16)) +
+  geom_bar()
+
+reduced_data %>%
+  ggplot(aes(x = nation_race, fill = pres16)) +
+  geom_bar()
+
+reduced_data %>%
+  ggplot(aes(x = nation_arms, fill = pres16)) +
+  geom_bar()
+
+reduced_data %>%
+  ggplot(aes(x = nation_welfare, fill = pres16)) +
+  geom_bar()
+
+reduced_data %>%
+  ggplot(aes(x = Assistance_to_the_poor, fill = pres16)) +
+  geom_bar()
+
+reduced_data %>%
+  ggplot(aes(x = Assistance_to_other_countries, fill = pres16)) +
+  geom_bar()
+
+reduced_data %>%
+  ggplot(aes(x = social_security, fill = pres16)) +
+  geom_bar()
+
+reduced_data %>%
+  ggplot(aes(x = race, fill = pres16)) +
+  geom_bar()
+```
 

### 2022-03-18 outputs/paper/paper.Rmd
ggplot
adeded age data

@@
 library(ggplot2)
+reduced_data %>%
+  ggplot(aes(x = age)) +
+  geom_bar()
+
+reduced_data %>%
+  ggplot(aes(x = gender)) +
+  geom_bar()
+
+reduced_data %>%
+  ggplot(aes(x = race)) +
+  geom_bar()
+
 reduced_data %>%

### 2022-03-18 scripts/01-data-preparation.R

@@ -21,7 +21,8 @@ names(raw_data)
 
 reduced_data <- 
   raw_data %>% 
-  select(sexnow1, 
+  select(age,
+         sexnow1, 
          degree,
          born,
          degree,

### 2022-03-18 outputs/paper/paper.Rmd
interpretation (results)

@@
 reduced_data %>%
-  ggplot(aes(x = age)) +
+  ggplot(aes(x = pres16)) +
   geom_bar()
@@
   geom_bar()
+
+reduced_data %>%
+  ggplot(aes(x = degree, fill = Family_income)) +
+  geom_bar()
+
 ```
+US demographics: White 76.3%, Black: 13.4%, American Indian: 1.3%, Asian: 5.9%, Hispanic: 18.5%
+https://www.census.gov/quickfacts/fact/table/US/PST045221.
+Considering the result above, our data sample is fairly recruited.
+
+1. nation_race 
+-> This session clearly shows the difference in approval rating between Hillary Clinton and 
+Donald Trump. Majority of people who thinks that national spending on improving conditions on
+race has to be increased advocated Hillary Clinton. On the other hand, people who did not agree 
+with them advocated Donald Trump. I assume that this is due to the image of Donald Trump.
+
+2. nation_arms
+-> Interestingly this session shows opposite trend compared to nation_race. Aside from people who
+answered "About right" and "NA", the two bars, "Too little" and "Too much" shows us opposite trend.
+
+In general, similar to the indicators above, Hillary Clinton supporters were moved by social equality such as 
+"Assistance to the Poor", "nation welfare", "nation education" and "Assistance to the other countries".
+On the other hand, Donald Trump supporters were interested by more conservative ideas that US is spending too much on the sessions above.
+
+3.
 

### 2022-03-18 outputs/paper/paper.Rmd
..
ggplot cleaning

@@
 library(ggplot2)
+library(dplyr)
 reduced_data %>%
-  ggplot(aes(x = pres16)) +
-  geom_bar()
+  ggplot(aes(x = age, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  labs(fil = "Candidate")
+
+reduced_data$pres16 <- factor(reduced_data$pres16, levels = c("Clinton","Trump","Other","Did not vote"))
+reduced_data %>%
+  ggplot(aes(x = pres16, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  labs(fil = "Candidate")
 
 reduced_data %>%
-  ggplot(aes(x = gender)) +
-  geom_bar()
+  ggplot(aes(x = gender, fill = gender)) +
+  geom_bar() +
+  scale_fill_brewer(palette = "Blues")
 
 reduced_data %>%
-  ggplot(aes(x = race)) +
-  geom_bar()
+  ggplot(aes(x = race, fill = race)) +
+  geom_bar() +
+  scale_fill_brewer(palette = "yellow")
 
 reduced_data %>%
-  ggplot(aes(x = gender, fill = pres16)) +
-  geom_bar()
+  ggplot(aes(x = gender, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  labs(fil = "Candidate")
 
 reduced_data %>%
-  ggplot(aes(x = degree, fill = pres16)) +
-  geom_bar()
+  ggplot(aes(x = degree, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  labs(fil = "Candidate")
 
 reduced_data %>%
-  ggplot(aes(x = born_US, fill = pres16)) +
-  geom_bar()
+  ggplot(aes(x = born_US, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  labs(fil = "Candidate")
+
+reduced_data$Family_income <- factor(reduced_data$Family_income, levels = c("Under $10,000","$10,000 ~ $19,999","$20,000 ~ $39,999","$40,000 ~ $59,999","$60,000 ~ $89,999","$90,000 ~ $129,999","$130,000 ~ $169,999","$170,000 or over"))
+reduced_data %>%
+  ggplot(aes(x = Family_income, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  labs(fil = "Candidate") +
+  coord_flip()
 
 reduced_data %>%
-  ggplot(aes(x = Family_income, fill = pres16)) +
-  geom_bar()
+  ggplot(aes(x = nation_education, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  labs(fil = "Candidate")
 
 reduced_data %>%
-  ggplot(aes(x = nation_education, fill = pres16)) +
-  geom_bar()
+  ggplot(aes(x = nation_race, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  labs(fil = "Candidate")
 
 reduced_data %>%
-  ggplot(aes(x = nation_race, fill = pres16)) +
-  geom_bar()
+  ggplot(aes(x = nation_arms, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  labs(fil = "Candidate")
 
 reduced_data %>%
-  ggplot(aes(x = nation_arms, fill = pres16)) +
-  geom_bar()
+  ggplot(aes(x = nation_welfare, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  labs(fil = "Candidate")
 
 reduced_data %>%
-  ggplot(aes(x = nation_welfare, fill = pres16)) +
-  geom_bar()
+  ggplot(aes(x = Assistance_to_the_poor, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  labs(fil = "Candidate")
 
 reduced_data %>%
-  ggplot(aes(x = Assistance_to_the_poor, fill = pres16)) +
-  geom_bar()
+  ggplot(aes(x = Assistance_to_other_countries, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  labs(fil = "Candidate")
 
 reduced_data %>%
-  ggplot(aes(x = Assistance_to_other_countries, fill = pres16)) +
-  geom_bar()
+  ggplot(aes(x = social_security, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  labs(fil = "Candidate")
 
 reduced_data %>%
-  ggplot(aes(x = social_security, fill = pres16)) +
-  geom_bar()
+  ggplot(aes(x = race, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  labs(fil = "Candidate")
 
 reduced_data %>%
-  ggplot(aes(x = race, fill = pres16)) +
-  geom_bar()
-
-reduced_data %>%
-  ggplot(aes(x = degree, fill = Family_income)) +
-  geom_bar()
+  ggplot(aes(x = degree, fill = factor(Family_income, levels = c("$170,000 or over","$130,000 ~ $169,999","$90,000 ~ $129,999","$60,000 ~ $89,999","$40,000 ~ $59,999","$20,000 ~ $39,999","$10,000 ~ $19,999","Under $10,000")))) +
+  geom_bar() +
+  labs(fill = "Family Income") +
+  coord_flip()
 
@@
 On the other hand, Donald Trump supporters were interested by more conservative ideas that US is spending too much on the sessions above.
-
-3.
 

### 2022-03-18 outputs/paper/paper.Rmd
gg plot and added tax

@@
   geom_bar() +
-  scale_fill_brewer(palette = "Blues")
+  scale_fill_brewer(palette = "pink")
 
@@
   geom_bar() +
-  scale_fill_brewer(palette = "yellow")
+  scale_fill_brewer(palette = "green")
 
@@
 
+reduced_data$tax <- factor(reduced_data$tax, levels = c("Too low","About right","Too high"))
+reduced_data %>%
+  ggplot(aes(x = tax, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  labs(fil = "Candidate")
+
 ```
@@
 
-In general, similar to the indicators above, Hillary Clinton supporters were moved by social equality such as 
+In general, similar to the indicators above, Hillary Clinton supporters were moved by improvement in social equality or welfare such as 
 "Assistance to the Poor", "nation welfare", "nation education" and "Assistance to the other countries".

### 2022-03-18 scripts/01-data-preparation.R

@@ -46,7 +46,8 @@ reduced_data <-
          natfarey,
          nataidy,
          natsoc,
-         pres16) %>% 
+         pres16,
+         tax) %>% 
   rename(gender = sexnow1,
          born_US = born,
          Family_income = income16,
@@ -259,6 +260,16 @@ reduced_data <-
     pres16 == 2 ~ "Trump",
     pres16 == 3 ~ "Other",
     pres16 == 4 ~ "Did not vote",
+  ))
+# Recode to rename variables in tax according to options from the Codebook 'GSS 2021 Codebook R1b.pdf'.
+#The question was do you consider the amount of federal income tax which you have to pay as too high
+#about right, or too low?
+reduced_data <- 
+  reduced_data %>% 
+  mutate(tax = case_when(
+    tax == 1 ~ "Too high",
+    tax == 2 ~ "About right",
+    tax == 3 ~ "Too low",
   ))
 
 reduced_data <- reduced_data[!is.na(reduced_data$pres16), ]

### 2022-03-19 scripts/01-data-preparation.R
gg plot final

@@ -47,7 +47,9 @@ reduced_data <-
          nataidy,
          natsoc,
          pres16,
-         tax) %>% 
+         tax,
+         immlimit,
+         ) %>% 
   rename(gender = sexnow1,
          born_US = born,
          Family_income = income16,
@@ -57,10 +59,21 @@ reduced_data <-
          nation_welfare = natfare,
          Assistance_to_the_poor = natfarey,
          Assistance_to_other_countries = nataidy,
-         social_security = natsoc)
+         social_security = natsoc,
+         Immigration = immlimit)
 
 rm(raw_data)
-
+# # Recode to rename variables in imigration according to options from the Codebook 'GSS 2021 Codebook R1b.pdf'.
+# The question was America should limit immigration in order to protect our national way of life
+reduced_data <- 
+  reduced_data %>% 
+  mutate(Immigration = case_when(
+    Immigration == 1 ~ "Strongly agree",
+    Immigration == 2 ~ "Agree",
+    Immigration == 3 ~ "Neither",
+    Immigration == 4 ~ "Disagree",
+    Immigration == 5 ~ "Strongly disagree"
+  ))
 # Recode to rename variables in gender according to options from the Codebook 'GSS 2021 Codebook R1b.pdf'.
 # The question was Do you describe yourself as male, female, or transgender?
 reduced_data <- 
@@ -260,16 +273,6 @@ reduced_data <-
     pres16 == 2 ~ "Trump",
     pres16 == 3 ~ "Other",
     pres16 == 4 ~ "Did not vote",
-  ))
-# Recode to rename variables in tax according to options from the Codebook 'GSS 2021 Codebook R1b.pdf'.
-#The question was do you consider the amount of federal income tax which you have to pay as too high
-#about right, or too low?
-reduced_data <- 
-  reduced_data %>% 
-  mutate(tax = case_when(
-    tax == 1 ~ "Too high",
-    tax == 2 ~ "About right",
-    tax == 3 ~ "Too low",
   ))
 
 reduced_data <- reduced_data[!is.na(reduced_data$pres16), ]

### 2022-03-19 outputs/paper/paper.Rmd
ggplot and glimpse of writing

@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "First sentence (what was done). Second sentence(what was found). Third sentence (why this matters). Fourth sentence."
 output:
@@
 # Introduction
-
+1) broader context to motivate
+2) some detail about hat the papaer is about
+3) a clear gap that needs to be filled
+4) what was done
+5) what was found
+6) why it is important
+7) The structure of the paper
+(3~4 paragraphs)
 Donald Trump won the 2016 United States presidential election. We are not US citizens. 
@@
 # Data
-1.Conveying a sense of the dataset of interest.
-2.Dicussion of the survey's methodology and its key features, strengths, weaknesses.
-For instance what is the population, frame and sample.
-How is the sample recruited 
-what sampling approach is taken
-what are some of the trade offs of this 
-how is non response handled.
-(What is good and bad about it)
+A sense of the datasets should be communicated to the reader.
+All varaiables should be throughly examined and explained.
+Explain if there were similar datasets that could have been used and why they were not.
+Why did we clean some variables?
+what the variables look like using graphs or tables discussion of the observations.
+Summary statistics should also be included
 
@@
 
-reduced_data$tax <- factor(reduced_data$tax, levels = c("Too low","About right","Too high"))
+reduced_data$Immigration <- factor(reduced_data$Immigration, levels = c("Strongly disagree","Disagree", "Neither","Agree","Strongly agree"))
 reduced_data %>%
-  ggplot(aes(x = tax, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  ggplot(aes(x = Immigration, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
   geom_bar() +
   scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
-  labs(fil = "Candidate")
+  labs(fil = "Candidate") +
+  coord_flip()
 
 ```
+
+# Results
+Summary statistics, tables, graphs, images .... 
+Regression tables must not contain stars??
+
+Thesis: what triggered U.S citizens to vote Clinton and Trump.
+
 US demographics: White 76.3%, Black: 13.4%, American Indian: 1.3%, Asian: 5.9%, Hispanic: 18.5%
@@
 
-# Results
-Convey findings.
-
 # Discussion
 
-Things that we must include
+what is done in this paper?
+what is something that we learn about the world
+what is another thing that we learn about the world
+what are some weaknesses of what was done?
+what is left to learn or how should we proceed in the future?
 

### 2022-03-19 outputs/paper/paper.Rmd
rmd knittable

--- 2022-03-18 outputs/paper/paper.Rmd
@@
 ---
-title: "Numerous variables relating to 2016 election in 2021 US GSS"
+title: "Numerous variables relating to 2016 election in 2021 US GSS^[Code and data are available at: https://github.com/SANGWOONG-LEE/2016electionanalysis.git]"
 subtitle: "Possible factors effecting 2016 election"
@@
   - Young Suk
-thanks: "Code and data are available at: https://github.com/SANGWOONG-LEE/2016electionanalysis.git".
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
 
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
+library(tidyverse)
+library(dplyr)
+```
+
 # Introduction
@@
 # Data
+```{r, echo=FALSE}
+reduced_data <- read_csv("outputs/data/prepared_gss.csv")
+```
 1.Conveying a sense of the dataset of interest.
@@
 
-```{r}
+```{r, echo=FALSE}
 library(ggplot2)
-library(dplyr)
 reduced_data %>%
-  ggplot(aes(x = age, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
-  labs(fil = "Candidate")
-
-reduced_data$pres16 <- factor(reduced_data$pres16, levels = c("Clinton","Trump","Other","Did not vote"))
-reduced_data %>%
-  ggplot(aes(x = pres16, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
-  labs(fil = "Candidate")
+  ggplot(aes(x = age)) +
+  geom_bar(na.rm = T)
 
 reduced_data %>%
-  ggplot(aes(x = gender, fill = gender)) +
-  geom_bar() +
-  scale_fill_brewer(palette = "Blues")
+  ggplot(aes(x = gender)) +
+  geom_bar()
 
 reduced_data %>%
-  ggplot(aes(x = race, fill = race)) +
-  geom_bar() +
-  scale_fill_brewer(palette = "yellow")
+  ggplot(aes(x = race)) +
+  geom_bar()
 
 reduced_data %>%
-  ggplot(aes(x = gender, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
-  labs(fil = "Candidate")
+  ggplot(aes(x = gender, fill = pres16)) +
+  geom_bar()
 
 reduced_data %>%
-  ggplot(aes(x = degree, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
-  labs(fil = "Candidate")
+  ggplot(aes(x = degree, fill = pres16)) +
+  geom_bar()
 
 reduced_data %>%
-  ggplot(aes(x = born_US, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
-  labs(fil = "Candidate")
-
-reduced_data$Family_income <- factor(reduced_data$Family_income, levels = c("Under $10,000","$10,000 ~ $19,999","$20,000 ~ $39,999","$40,000 ~ $59,999","$60,000 ~ $89,999","$90,000 ~ $129,999","$130,000 ~ $169,999","$170,000 or over"))
-reduced_data %>%
-  ggplot(aes(x = Family_income, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
-  labs(fil = "Candidate") +
-  coord_flip()
+  ggplot(aes(x = born_US, fill = pres16)) +
+  geom_bar()
 
 reduced_data %>%
-  ggplot(aes(x = nation_education, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
-  labs(fil = "Candidate")
+  ggplot(aes(x = Family_income, fill = pres16)) +
+  geom_bar()  + coord_flip()
 
 reduced_data %>%
-  ggplot(aes(x = nation_race, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
-  labs(fil = "Candidate")
+  ggplot(aes(x = nation_education, fill = pres16)) +
+  geom_bar()
 
 reduced_data %>%
-  ggplot(aes(x = nation_arms, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
-  labs(fil = "Candidate")
+  ggplot(aes(x = nation_race, fill = pres16)) +
+  geom_bar()
 
 reduced_data %>%
-  ggplot(aes(x = nation_welfare, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
-  labs(fil = "Candidate")
+  ggplot(aes(x = nation_arms, fill = pres16)) +
+  geom_bar()
 
 reduced_data %>%
-  ggplot(aes(x = Assistance_to_the_poor, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
-  labs(fil = "Candidate")
+  ggplot(aes(x = nation_welfare, fill = pres16)) +
+  geom_bar()
 
 reduced_data %>%
-  ggplot(aes(x = Assistance_to_other_countries, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
-  labs(fil = "Candidate")
+  ggplot(aes(x = Assistance_to_the_poor, fill = pres16)) +
+  geom_bar()
 
 reduced_data %>%
-  ggplot(aes(x = social_security, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
-  labs(fil = "Candidate")
+  ggplot(aes(x = Assistance_to_other_countries, fill = pres16)) +
+  geom_bar()
 
 reduced_data %>%
-  ggplot(aes(x = race, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
-  labs(fil = "Candidate")
+  ggplot(aes(x = social_security, fill = pres16)) +
+  geom_bar()
 
 reduced_data %>%
-  ggplot(aes(x = degree, fill = factor(Family_income, levels = c("$170,000 or over","$130,000 ~ $169,999","$90,000 ~ $129,999","$60,000 ~ $89,999","$40,000 ~ $59,999","$20,000 ~ $39,999","$10,000 ~ $19,999","Under $10,000")))) +
-  geom_bar() +
-  labs(fill = "Family Income") +
-  coord_flip()
-
+  ggplot(aes(x = race, fill = pres16)) +
+  geom_bar()
 ```
-US demographics: White 76.3%, Black: 13.4%, American Indian: 1.3%, Asian: 5.9%, Hispanic: 18.5%
-https://www.census.gov/quickfacts/fact/table/US/PST045221.
-Considering the result above, our data sample is fairly recruited.
-
-1. nation_race 
--> This session clearly shows the difference in approval rating between Hillary Clinton and 
-Donald Trump. Majority of people who thinks that national spending on improving conditions on
-race has to be increased advocated Hillary Clinton. On the other hand, people who did not agree 
-with them advocated Donald Trump. I assume that this is due to the image of Donald Trump.
-
-2. nation_arms
--> Interestingly this session shows opposite trend compared to nation_race. Aside from people who
-answered "About right" and "NA", the two bars, "Too little" and "Too much" shows us opposite trend.
-
-In general, similar to the indicators above, Hillary Clinton supporters were moved by social equality such as 
-"Assistance to the Poor", "nation welfare", "nation education" and "Assistance to the other countries".
-On the other hand, Donald Trump supporters were interested by more conservative ideas that US is spending too much on the sessions above.
 

### 2022-03-19 outputs/paper/paper.Rmd

--- 2022-03-19 outputs/paper/paper.Rmd
@@
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
+library(tidyverse)
+library(dplyr)
+```
 
@@
 # Data
+```{r, include=FALSE, echo=FALSE}
+reduced_data <- read_csv("outputs/data/prepared_gss.csv")
+```
+1.Conveying a sense of the dataset of interest.
+2.Dicussion of the survey's methodology and its key features, strengths, weaknesses.
+For instance what is the population, frame and sample.
+How is the sample recruited 
+what sampling approach is taken
+what are some of the trade offs of this 
+how is non response handled.
+(What is good and bad about it
+
 A sense of the datasets should be communicated to the reader.
@@
 
-```{r}
-library(ggplot2)
-library(dplyr)
+```{r, echo=FALSE}
+ggplot(reduced_data, aes(x = age)) +
+  geom_bar(na.rm = T)
+
 reduced_data %>%
@@
 reduced_data$pres16 <- factor(reduced_data$pres16, levels = c("Clinton","Trump","Other","Did not vote"))
+
 reduced_data %>%
@@
 reduced_data %>%
-  ggplot(aes(x = degree, fill = factor(Family_income, levels = c("$170,000 or over","$130,000 ~ $169,999","$90,000 ~ $129,999","$60,000 ~ $89,999","$40,000 ~ $59,999","$20,000 ~ $39,999","$10,000 ~ $19,999","Under $10,000")))) +
+  ggplot(aes(x = degree, fill = factor(Family_income, 
+                                       levels = c("$170,000 or over","$130,000 ~ $169,999","$90,000 ~ $129,999","$60,000 ~ $89,999",
+                                                  "$40,000 ~ $59,999","$20,000 ~ $39,999","$10,000 ~ $19,999","Under $10,000")))) +
   geom_bar() +
@@
 reduced_data$Immigration <- factor(reduced_data$Immigration, levels = c("Strongly disagree","Disagree", "Neither","Agree","Strongly agree"))
+
 reduced_data %>%

### 2022-03-19 outputs/paper/paper.Rmd
brief introduction

--- 2022-03-19 outputs/paper/paper.Rmd
@@
 Summary statistics should also be included
+
+Original dataset is collected through US GSS, which is aimed for the goal that enabling the cleanest comparison to show changes in public opinion in the United States.
+Usually, it is conducted in person, however because of pandemic, it was conducted via online.
+Among numerous interesting findings from the dataset, we choose to analyze the 2016 United States presidential election.
+However, choosing specific variables related to the election was not easy since deciding who to vote is followed by countless factors.After looking over the variables, we chose age,race,gender,degree,born US, Family income, national spending on education, race, arms, welfare, the poor, the other countries, social security and immigration as potential variables.
+We mainly divided the variables into two parts: personal information and thoughts on national policy.
+Age, race, gender, degree, born US and family income are respondents' personal information, and national spending on education, race, arms, welfare, the poor, the other countries, social security and immigration are personal thoughts on national policy.
+
+US demographics: White 76.3%, Black: 13.4%, American Indian: 1.3%, Asian: 5.9%, Hispanic: 18.5%
+https://www.census.gov/quickfacts/fact/table/US/PST045221.
+Considering the result above, our data sample is fairly recruited.
 
@@
 
-US demographics: White 76.3%, Black: 13.4%, American Indian: 1.3%, Asian: 5.9%, Hispanic: 18.5%
-https://www.census.gov/quickfacts/fact/table/US/PST045221.
-Considering the result above, our data sample is fairly recruited.
-
 1. nation_race 

### 2022-03-19 outputs/paper/paper.Rmd
ggplot

--- 2022-03-19 outputs/paper/paper.Rmd
@@
 ```
+The data comes from US General Social Survey.
 1.Conveying a sense of the dataset of interest.
@@
 ```{r, echo=FALSE}
-ggplot(reduced_data, aes(x = age)) +
-  geom_bar(na.rm = T)
+reduced_data %>%
+  drop_na(age_group) %>%
+  ggplot(aes(x = age_group, fill = age_group)) +
+  geom_bar() +
+  coord_flip() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Survey Respondents' Age",
+       caption = "Figure 1: Distribution of survey respondents' age group") +
+  theme(legend.position = "none")
 
 reduced_data %>%
-  ggplot(aes(x = age, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  ggplot(aes(x = race, fill = race)) +
   geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
-  labs(fil = "Candidate")
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Survey Respondents' race",
+       caption = "Figure 2: Distribution of survey respondents' Race") +
+  theme(legend.position = "none")
+
+# reduced_data %>%
+#   ggplot(aes(x = age, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+#   geom_bar() +
+#   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
+#   labs(fil = "Candidate")
 
@@
 reduced_data %>%
+  drop_na(nation_race) %>%
+  ggplot(aes(x = race, fill = nation_race)) +
+  geom_bar() +
+  coord_flip() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Opinion on National Spending on Improving Conditions of Black Population",
+       caption = "Figure 3: Distribution of survey respondents' Race") +
+  facet_wrap(~ pres16)
+
+reduced_data %>%
   ggplot(aes(x = pres16, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
   geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
   labs(fil = "Candidate")
 
-reduced_data %>%
-  ggplot(aes(x = gender, fill = gender)) +
-  geom_bar() +
-  scale_fill_brewer(palette = "pink")
-
-reduced_data %>%
-  ggplot(aes(x = race, fill = race)) +
-  geom_bar() +
-  scale_fill_brewer(palette = "green")
+# reduced_data %>%
+#   ggplot(aes(x = gender, fill = gender)) +
+#   geom_bar() +
+#   scale_fill_brewer(palette = "pink")
 
@@
   geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
   labs(fil = "Candidate")
@@
   geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
   labs(fil = "Candidate")
 
-reduced_data$Family_income <- factor(reduced_data$Family_income, levels = c("Under $10,000","$10,000 ~ $19,999","$20,000 ~ $39,999","$40,000 ~ $59,999","$60,000 ~ $89,999","$90,000 ~ $129,999","$130,000 ~ $169,999","$170,000 or over"))
+reduced_data$Family_income <- factor(reduced_data$Family_income, 
+                                     levels = c("Under $10,000","$10,000 ~ $19,999","$20,000 ~ $39,999","$40,000 ~ $59,999", 
+                                                "$60,000 ~ $89,999","$90,000 ~ $129,999","$130,000 ~ $169,999","$170,000 or over"))
 reduced_data %>%
@@
   geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
+  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
   labs(fil = "Candidate") +

### 2022-03-19 scripts/01-data-preparation.R

@@ -63,6 +63,20 @@ reduced_data <-
          Immigration = immlimit)
 
 rm(raw_data)
+
+# Add age_group column based on age column
+reduced_data <- reduced_data %>% 
+  mutate(age_group = reduced_data$age %>% cut(breaks = seq(19, 89, by = 10))) %>%
+  mutate(age_group = case_when(
+    age_group == "(19,29]" ~ "20-29 years old",
+    age_group == "(29,39]" ~ "30-39 years old",
+    age_group == "(39,49]" ~ "40-49 years old",
+    age_group == "(49,59]" ~ "50-59 years old",
+    age_group == "(59,69]" ~ "60-69 years old",
+    age_group == "(69,79]" ~ "70-79 years old",
+    age_group == "(79,89]" ~ "80-89 years old"
+  ))
+
 # # Recode to rename variables in imigration according to options from the Codebook 'GSS 2021 Codebook R1b.pdf'.
 # The question was America should limit immigration in order to protect our national way of life
 reduced_data <- 
@@ -192,18 +206,18 @@ reduced_data$race_sum <- reduced_data %>%
 # Create a column named "race" and assign values according to the columns from the survey.
 reduced_data <- reduced_data %>%
   mutate(race = case_when(
-    raceacs1 == 1 ~ "white",
-    raceacs2 == 1 ~ "black",
-    raceacs3 == 1 ~ "american_native",
-    raceacs4 == 1 ~ "asian",
-    raceacs5 == 1 ~ "asian",
-    raceacs6 == 1 ~ "asian",
-    raceacs7 == 1 ~ "asian",
-    raceacs8 == 1 ~ "asian",
-    raceacs9 == 1 ~ "asian",
-    raceacs10 == 1 ~ "asian",
-    raceacs15 == 1 ~ "other",
-    raceacs16 == 1 ~ "hispanic"
+    raceacs1 == 1 ~ "White",
+    raceacs2 == 1 ~ "Black",
+    raceacs3 == 1 ~ "American Native",
+    raceacs4 == 1 ~ "Asian",
+    raceacs5 == 1 ~ "Asian",
+    raceacs6 == 1 ~ "Asian",
+    raceacs7 == 1 ~ "Asian",
+    raceacs8 == 1 ~ "Asian",
+    raceacs9 == 1 ~ "Asian",
+    raceacs10 == 1 ~ "Asian",
+    raceacs15 == 1 ~ "Other",
+    raceacs16 == 1 ~ "Hispanic"
[…]

### 2022-03-19 outputs/paper/paper.Rmd
ggplot

@@
 ```
-The data comes from US General Social Survey.
+* number of survey participants of age in the range of 20~40 is unexpectedly low -> not many people in ages between 20-40 participated in the survey
+* Clinton's vote is higher than trump's vote in this survey, which contradicts the election result of 2016, reader should keep in mind that this survey is more biased towards clinton.
+
 1.Conveying a sense of the dataset of interest.
@@
 
+reduced_data %>%
+  ggplot(aes(x = pres16, fill = pres16)) +
+  geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Surveyee's vote choice of 2016 Presidential election",
+       caption = "Figure 3: Distribution of survey respondents' vote choice") +
+  theme(legend.position = "none")
+
 # reduced_data %>%
@@
   drop_na(nation_race) %>%
+  drop_na(race) %>%
   ggplot(aes(x = race, fill = nation_race)) +
@@
        title = "Opinion on National Spending on Improving Conditions of Black Population",
-       caption = "Figure 3: Distribution of survey respondents' Race") +
-  facet_wrap(~ pres16)
+       caption = "Figure 4: Distribution of survey respondents' Race") +
+  facet_wrap(~ pres16) + 
+  theme(legend.title = element_blank())
 

### 2022-03-19 outputs/paper/paper.Rmd
ggplot

@@
 ---
-title: "Numerous variables relating to 2016 election in 2021 US GSS^[Code and data are available at: https://github.com/SANGWOONG-LEE/2016electionanalysis.git]"
+title: "Numerous variables relating to 2016 election in 2021 US GSS^[Code and data are available at: https://github.com/SANGWOONG-LEE/2016electionanalysis]"
 subtitle: "Possible factors effecting 2016 election"
@@
 ```{r, echo=FALSE}
+reduced_data$pres16 <- factor(reduced_data$pres16, levels = c("Clinton","Trump","Other","Did not vote"))
+
+reduced_data$Family_income <- factor(reduced_data$Family_income, 
+                                     levels = c("Under $10,000","$10,000 ~ $19,999","$20,000 ~ $39,999","$40,000 ~ $59,999", 
+                                                "$60,000 ~ $89,999","$90,000 ~ $129,999","$130,000 ~ $169,999","$170,000 or over"))
+
+reduced_data$Immigration <- factor(reduced_data$Immigration, levels = c("Strongly disagree","Disagree", "Neither","Agree","Strongly agree"))
+
 reduced_data %>%
@@
 reduced_data %>%
-  ggplot(aes(x = pres16, fill = pres16)) +
+  ggplot(aes(x = pres16, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
   geom_bar() +
@@
        y = element_blank(),
-       title = "Surveyee's vote choice of 2016 Presidential election",
+       title = "Surveyee's Vote Choice of 2016 Presidential Election",
        caption = "Figure 3: Distribution of survey respondents' vote choice") +
   theme(legend.position = "none")
-
-# reduced_data %>%
-#   ggplot(aes(x = age, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-#   geom_bar() +
-#   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-#   labs(fil = "Candidate")
-
-reduced_data$pres16 <- factor(reduced_data$pres16, levels = c("Clinton","Trump","Other","Did not vote"))
 
@@
 reduced_data %>%
-  ggplot(aes(x = pres16, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  ggplot(aes(x = born_US, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
   geom_bar() +
   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate")
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Surveyee's Vote Choice of 2016 Presidential Election by Brith Place",
+       caption = "Figure 5: Distribution of survey respondents' vote choice by birth place",
+       fil = "Candidate")
 
@@
 
-reduced_data %>%
-  ggplot(aes(x = gender, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate")
+# reduced_data %>%
+#   ggplot(aes(x = gender, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+#   geom_bar() +
+#   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
+#   labs(fil = "Candidate")
 
@@
 
-reduced_data %>%
-  ggplot(aes(x = born_US, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate")
-
-reduced_data$Family_income <- factor(reduced_data$Family_income, 
-                                     levels = c("Under $10,000","$10,000 ~ $19,999","$20,000 ~ $39,999","$40,000 ~ $59,999", 
-                                                "$60,000 ~ $89,999","$90,000 ~ $129,999","$130,000 ~ $169,999","$170,000 or over"))
 reduced_data %>%
@@
   coord_flip()
-
-reduced_data$Immigration <- factor(reduced_data$Immigration, levels = c("Strongly disagree","Disagree", "Neither","Agree","Strongly agree"))
 

### 2022-03-19 scripts/01-data-preparation.R

@@ -222,7 +222,7 @@ reduced_data <- reduced_data %>%
 
 # If a person is of mixed race, that person is assigned value "other"
 reduced_data <- reduced_data %>%
-  mutate(race = if_else(race_sum == 1, reduced_data$race, "other"))
+  mutate(race = if_else(race_sum == 1, reduced_data$race, "Other"))
 
 # Remove columns that are no longer needed.
 reduced_data <- reduced_data %>%

### 2022-03-19 outputs/paper/paper.Rmd
ggplot

@@
        y = element_blank(),
-       title = "Opinion on National Spending on Improving Conditions of Black Population",
-       caption = "Figure 4: Distribution of survey respondents' Race") +
+       title = "Is the US Governemnt Spending Enough to Improve Conditions of Black Population?",
+       caption = "Figure 4: Opinion on National Spending on Improving Conditions of Black Population") +
   facet_wrap(~ pres16) + 
@@
 #   labs(fil = "Candidate")
+
+reduced_data %>%
+  drop_na(nation_welfare) %>%
+  ggplot(aes(x = race, fill = nation_welfare)) +
+  geom_bar() +
+  coord_flip() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Is the US Government Spending Enough on Welfare?",
+       caption = "Figure 6: Opinion on National Spending on welfare") +
+  facet_wrap(~ pres16) + 
+  theme(legend.title = element_blank())
+```
+
+``````{r, echo=FALSE}
 
@@
 
-reduced_data %>%
-  ggplot(aes(x = nation_welfare, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate")
+# reduced_data %>%
+#   ggplot(aes(x = nation_welfare, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+#   geom_bar() +
+#   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
+#   labs(fil = "Candidate")
 

### 2022-03-19 outputs/paper/paper.Rmd
ggplot

@@
 reduced_data <- read_csv("outputs/data/prepared_gss.csv")
+reduced_data$pres16 <- factor(reduced_data$pres16, levels = c("Clinton","Trump","Other","Did not vote"))
+
+reduced_data$Family_income <- factor(reduced_data$Family_income, 
+                                     levels = c("Under $10,000","$10,000 ~ $19,999","$20,000 ~ $39,999","$40,000 ~ $59,999", 
+                                                "$60,000 ~ $89,999","$90,000 ~ $129,999","$130,000 ~ $169,999","$170,000 or over"))
+
+reduced_data$Immigration <- factor(reduced_data$Immigration, levels = c("Strongly disagree","Disagree", "Neither","Agree","Strongly agree"))
 ```
@@
 ```{r, echo=FALSE}
-reduced_data$pres16 <- factor(reduced_data$pres16, levels = c("Clinton","Trump","Other","Did not vote"))
-
-reduced_data$Family_income <- factor(reduced_data$Family_income, 
-                                     levels = c("Under $10,000","$10,000 ~ $19,999","$20,000 ~ $39,999","$40,000 ~ $59,999", 
-                                                "$60,000 ~ $89,999","$90,000 ~ $129,999","$130,000 ~ $169,999","$170,000 or over"))
-
-reduced_data$Immigration <- factor(reduced_data$Immigration, levels = c("Strongly disagree","Disagree", "Neither","Agree","Strongly agree"))
 
@@
 reduced_data %>%
+  ggplot(aes(x = gender, fill = gender)) +
+  geom_bar() +
+  #scale_fill_manual("gender", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Survey Respondents' Gender",
+       caption = "Figure 3: Distribution of survey respondents' gender") +
+  theme(legend.position = "none")
+
+reduced_data %>%
   ggplot(aes(x = pres16, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
@@
        title = "Surveyee's Vote Choice of 2016 Presidential Election",
-       caption = "Figure 3: Distribution of survey respondents' vote choice") +
+       caption = "Figure 4: Distribution of survey respondents' vote choice") +
   theme(legend.position = "none")
@@
        y = element_blank(),
-       title = "Is the US Governemnt Spending Enough to Improve Conditions of Black Population?",
-       caption = "Figure 4: Opinion on National Spending on Improving Conditions of Black Population") +
+       title = "Is the US Spending Enough to Improve Conditions of Black Population?",
+       caption = "Figure 5: Opinion on National Spending on Improving Conditions of Black Population") +
   facet_wrap(~ pres16) + 
@@
        title = "Surveyee's Vote Choice of 2016 Presidential Election by Brith Place",
-       caption = "Figure 5: Distribution of survey respondents' vote choice by birth place",
+       caption = "Figure 6: Distribution of survey respondents' vote choice by birth place",
        fil = "Candidate")
@@
   drop_na(nation_welfare) %>%
+  drop_na(race) %>%
   ggplot(aes(x = race, fill = nation_welfare)) +
@@
        y = element_blank(),
-       title = "Is the US Government Spending Enough on Welfare?",
-       caption = "Figure 6: Opinion on National Spending on welfare") +
+       title = "Is the US Spending Enough on Welfare?",
+       caption = "Figure 7: Opinion on National Spending on welfare") +
   facet_wrap(~ pres16) + 
   theme(legend.title = element_blank())
+
+reduced_data %>%
+  ggplot(aes(x = degree, fill = factor(Family_income, 
+                                       levels = c("$170,000 or over","$130,000 ~ $169,999","$90,000 ~ $129,999","$60,000 ~ $89,999",
+                                                  "$40,000 ~ $59,999","$20,000 ~ $39,999","$10,000 ~ $19,999","Under $10,000")))) +
+  geom_bar(position = "dodge") +
+  labs(fill = "Family Income") +
+  coord_flip()
 ```
@@
 
-reduced_data %>%
-  ggplot(aes(x = nation_race, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate")
+# reduced_data %>%
+#   ggplot(aes(x = nation_race, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+#   geom_bar() +
+#   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
+#   labs(fil = "Candidate")
 
@@
 
-reduced_data %>%
-  ggplot(aes(x = social_security, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate")
+# reduced_data %>%
+#   ggplot(aes(x = social_security, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+#   geom_bar() +
+#   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
+#   labs(fil = "Candidate")
 
-reduced_data %>%
-  ggplot(aes(x = race, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate")
-
-reduced_data %>%
-  ggplot(aes(x = degree, fill = factor(Family_income, 
-                                       levels = c("$170,000 or over","$130,000 ~ $169,999","$90,000 ~ $129,999","$60,000 ~ $89,999",
-                                                  "$40,000 ~ $59,999","$20,000 ~ $39,999","$10,000 ~ $19,999","Under $10,000")))) +
-  geom_bar() +
-  labs(fill = "Family Income") +
-  coord_flip()
+# reduced_data %>%
+#   ggplot(aes(x = race, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+#   geom_bar() +
+#   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
+#   labs(fil = "Candidate")
 
@@
   labs(fil = "Candidate") +
-  coord_flip()
+  coord_flip() +
+  facet_wrap(~born_US)
 

### 2022-03-19 outputs/paper/paper.Rmd
ggplot

@@
        y = element_blank(),
-       title = "Survey Respondents' race",
-       caption = "Figure 2: Distribution of survey respondents' Race") +
+       title = "Survey Respondents' Race",
+       caption = "Figure 2: Distribution of survey respondents' race") +
   theme(legend.position = "none")
@@
        title = "Is the US Spending Enough to Improve Conditions of Black Population?",
-       caption = "Figure 5: Opinion on National Spending on Improving Conditions of Black Population") +
+       caption = "Figure 5: Opinion on national spending on improving conditions of black population") +
   facet_wrap(~ pres16) + 
@@
        fil = "Candidate")
-
-# reduced_data %>%
-#   ggplot(aes(x = gender, fill = gender)) +
-#   geom_bar() +
-#   scale_fill_brewer(palette = "pink")
 
@@
        y = element_blank(),
-       title = "Is the US Spending Enough on Welfare?",
-       caption = "Figure 7: Opinion on National Spending on welfare") +
+       title = "Is the Government Spending Enough on Welfare?",
+       caption = "Figure 7: Opinion on national spending on welfare") +
   facet_wrap(~ pres16) + 
@@
 reduced_data %>%
+  drop_na(degree) %>%
   ggplot(aes(x = degree, fill = factor(Family_income, 
@@
   labs(fill = "Family Income") +
-  coord_flip()
-```
-
-``````{r, echo=FALSE}
+  coord_flip() +
+  facet_wrap(~pres16) +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Education, Income and its Effect to the Vote",
+       caption = "Figure 8: Distribution of votes by degree and family income")
 
 reduced_data %>%
-  ggplot(aes(x = degree, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate")
-
-reduced_data %>%
-  ggplot(aes(x = Family_income, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  ggplot(aes(x = Immigration, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
   geom_bar() +
@@
   labs(fil = "Candidate") +
-  coord_flip()
+  coord_flip() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Should the Government Limit Immigration?",
+       caption = "Figure 9: Opinion on wheter the US should limit immigration to protect their culture") +
+  facet_wrap(~pres16)
 
@@
   ggplot(aes(x = nation_education, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
+  geom_bar(position = "dodge") +
   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate")
+  labs(fil = "Candidate") +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Is the Government Spending Enough on Improving Education?",
+       caption = "Figure 10: Surveyee's opinion on government's budget on education improvement")
+
+reduced_data %>%
+  ggplot(aes(x = nation_arms, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar(position = "dodge") +
+  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
+  labs(fil = "Candidate") +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Is the Government Spending Enough on Arms and National Defense?",
+       caption = "Figure 11: Surveyee's opinion on government's budget on arms")
+
+reduced_data %>%
+  ggplot(aes(x = Assistance_to_the_poor, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar(position = "dodge") +
+  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
+  labs(fil = "Candidate") +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Is the Government Spending Enough to Assis the Poor?",
+       caption = "Figure 12: Surveyee's opinion on government's budget on assisting the poor")
+
+reduced_data %>%
+  ggplot(aes(x = Assistance_to_other_countries, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  geom_bar(position = "dodge") +
+  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
+  labs(fil = "Candidate") +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Is the Government Spending Enough to Assis Other Countries?",
+       caption = "Figure 13: Surveyee's opinion on government's budget on assisting other countries")
+```
+
+```{r, echo=FALSE}
+
+# reduced_data %>%
+#   ggplot(aes(x = degree, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+#   geom_bar() +
+#   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
+#   labs(fil = "Candidate")
+
+# reduced_data %>%
+#   ggplot(aes(x = Family_income, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+#   geom_bar() +
+#   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
+#   labs(fil = "Candidate") +
+#   coord_flip()
 
@@
 
-reduced_data %>%
-  ggplot(aes(x = nation_arms, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate")
-
 # reduced_data %>%
@@
 #   labs(fil = "Candidate")
-
-reduced_data %>%
-  ggplot(aes(x = Assistance_to_the_poor, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate")
-
-reduced_data %>%
-  ggplot(aes(x = Assistance_to_other_countries, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate")
 
@@
 #   labs(fil = "Candidate")
-
-reduced_data %>%
-  ggplot(aes(x = Immigration, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate") +
-  coord_flip() +
-  facet_wrap(~born_US)
 

### 2022-03-19 outputs/paper/paper.Rmd
ggplot

@@
 reduced_data %>%
+  ggplot(aes(x = pres16, fill = gender)) +
+  geom_bar() +
+  coord_flip() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Surveyee's Vote Choice of 2016 Presidential Election by Gender",
+       caption = "Figure 7: Distribution of survey respondents' vote by gender",)
+
+reduced_data %>%
   drop_na(nation_welfare) %>%
@@
        title = "Is the Government Spending Enough on Welfare?",
-       caption = "Figure 7: Opinion on national spending on welfare") +
+       caption = "Figure 8: Opinion on national spending on welfare") +
   facet_wrap(~ pres16) + 
@@
        title = "Education, Income and its Effect to the Vote",
-       caption = "Figure 8: Distribution of votes by degree and family income")
+       caption = "Figure 9: Distribution of votes by degree and family income")
 
@@
        title = "Should the Government Limit Immigration?",
-       caption = "Figure 9: Opinion on wheter the US should limit immigration to protect their culture") +
+       caption = "Figure 10: Opinion on wheter the US should limit immigration to protect their culture") +
   facet_wrap(~pres16)
@@
        title = "Is the Government Spending Enough on Improving Education?",
-       caption = "Figure 10: Surveyee's opinion on government's budget on education improvement")
+       caption = "Figure 11: Surveyee's opinion on government's budget on education improvement")
 
@@
        title = "Is the Government Spending Enough on Arms and National Defense?",
-       caption = "Figure 11: Surveyee's opinion on government's budget on arms")
+       caption = "Figure 12: Surveyee's opinion on government's budget on arms")
 
@@
        title = "Is the Government Spending Enough to Assis the Poor?",
-       caption = "Figure 12: Surveyee's opinion on government's budget on assisting the poor")
+       caption = "Figure 13: Surveyee's opinion on government's budget on assisting the poor")
 
@@
        title = "Is the Government Spending Enough to Assis Other Countries?",
-       caption = "Figure 13: Surveyee's opinion on government's budget on assisting other countries")
+       caption = "Figure 14: Surveyee's opinion on government's budget on assisting other countries")
 ```

### 2022-03-19 outputs/paper/paper.Rmd
data writing (1)

@@
 Among numerous interesting findings from the dataset, we choose to analyze the 2016 United States presidential election.
+This slightly deviates the US GSS goal, but we thought it clearly represents current situation of US.
 However, choosing specific variables related to the election was not easy since deciding who to vote is followed by countless factors.After looking over the variables, we chose age,race,gender,degree,born US, Family income, national spending on education, race, arms, welfare, the poor, the other countries, social security and immigration as potential variables.
-We mainly divided the variables into two parts: personal information and thoughts on national policy.
-Age, race, gender, degree, born US and family income are respondents' personal information, and national spending on education, race, arms, welfare, the poor, the other countries, social security and immigration are personal thoughts on national policy.
+We mainly divided the variables into two parts: personal information (Age, race, gender, degree, born US and family income) and thoughts on national policies (national spending on education, race, arms, welfare, the poor, the other countries, social security and immigration).
+As I mentioned above, deciding who to vote is followed by very complex system of thoughts. Thus, as we cut down the variables, they do not guarantee 100% of the vote.
 

### 2022-03-19 outputs/paper/paper.Rmd
ggplot

@@
 
-```{r, echo=FALSE}
+```{r fig1, echo=FALSE}
 
@@
   theme(legend.position = "none")
+```
+
+```{r fig2, echo=FALSE}
 
@@
   theme(legend.position = "none")
+```
 
+```{r fig3, echo=FALSE}
 reduced_data %>%
@@
   theme(legend.position = "none")
+```
 
+```{r fig4, echo=FALSE}
 reduced_data %>%
@@
   theme(legend.position = "none")
+```
 
+```{r fig5, echo=FALSE}
 reduced_data %>%
@@
   theme(legend.title = element_blank())
+```
 
+```{r fig6, echo=FALSE}
 reduced_data %>%
@@
        fil = "Candidate")
+```
 
-# reduced_data %>%
-#   ggplot(aes(x = gender, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-#   geom_bar() +
-#   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-#   labs(fil = "Candidate")
-
+```{r fig7, echo=FALSE}
 reduced_data %>%
@@
        caption = "Figure 7: Distribution of survey respondents' vote by gender",)
+```
 
+```{r fig8, echo=FALSE}
 reduced_data %>%
@@
   theme(legend.title = element_blank())
+```
 
+```{r fig9, echo=FALSE}
 reduced_data %>%
@@
        caption = "Figure 9: Distribution of votes by degree and family income")
+```
 
+```{r fig10, echo=FALSE}
 reduced_data %>%
@@
   facet_wrap(~pres16)
+```
 
+```{r fig11, echo=FALSE}
 reduced_data %>%
@@
        caption = "Figure 11: Surveyee's opinion on government's budget on education improvement")
+```
 
+```{r fig12, echo=FALSE}
 reduced_data %>%
@@
        caption = "Figure 12: Surveyee's opinion on government's budget on arms")
+```
 
+```{r fig13, echo=FALSE}
 reduced_data %>%
@@
        caption = "Figure 13: Surveyee's opinion on government's budget on assisting the poor")
+```
 
+```{r fig14, echo=FALSE}
 reduced_data %>%

### 2022-03-19 outputs/paper/paper.Rmd
figure and data sessions writing

@@
 Original dataset is collected through US GSS, which is aimed for the goal that enabling the cleanest comparison to show changes in public opinion in the United States.
-Usually, it is conducted in person, however because of pandemic, it was conducted via online.
+Usually, it is conducted in person, however because of pandemic, this time it was conducted via online.
 Among numerous interesting findings from the dataset, we choose to analyze the 2016 United States presidential election.
 This slightly deviates the US GSS goal, but we thought it clearly represents current situation of US.
-However, choosing specific variables related to the election was not easy since deciding who to vote is followed by countless factors.After looking over the variables, we chose age,race,gender,degree,born US, Family income, national spending on education, race, arms, welfare, the poor, the other countries, social security and immigration as potential variables.
+However, choosing specific variables related to the election was not easy since deciding who to vote is followed by countless factors.
+After looking over the variables, we chose age,race,gender,degree,born US, Family income, national spending on education, race, arms, welfare, the poor, the other countries, social security and immigration as potential variables.
 We mainly divided the variables into two parts: personal information (Age, race, gender, degree, born US and family income) and thoughts on national policies (national spending on education, race, arms, welfare, the poor, the other countries, social security and immigration).
-As I mentioned above, deciding who to vote is followed by very complex system of thoughts. Thus, as we cut down the variables, they do not guarantee 100% of the vote.
+As I mentioned above, deciding who to vote is followed by very complex system of thoughts.
+Thus, as we cut down the variables, they do not guarantee 100% of the vote.
+Also, because we did not exclude non-response from our dataset, for some graphs it caused difficulties in interpretation.
+There exists similar survey similar to ours, CNN exit polls, however it is not obtain the datasets they used.
+Now, let's get into the variables in details.
+One thing that we foucsed was whether respondents were equivalently recruited.
+To do this, we first have to look at respondents' personal information.
+If you take a look at fig(), there is not much to notice.
+Few things to notice are there are more female respondents than male, and for race, there are more White than others.
+However, following US demographics, White 76.3%, Black: 13.4%, American Indian: 1.3%, Asian: 5.9%, Hispanic: 18.5%, our sample is fairly well recruited.
 
-US demographics: White 76.3%, Black: 13.4%, American Indian: 1.3%, Asian: 5.9%, Hispanic: 18.5%
 https://www.census.gov/quickfacts/fact/table/US/PST045221.

### 2022-03-19 outputs/paper/paper.Rmd
Data writing done (1st)

@@
 output:
-  bookdown::pdf_document2
+  bookdown::pdf_document2:
+    fig_caption: yes
 toc: FALSE
@@
 library(dplyr)
+library(knitr)
 ```
@@
 ```
-* number of survey participants of age in the range of 20~40 is unexpectedly low -> not many people in ages between 20-40 participated in the survey
-* Clinton's vote is higher than trump's vote in this survey, which contradicts the election result of 2016, reader should keep in mind that this survey is more biased towards clinton.
-
-1.Conveying a sense of the dataset of interest.
-2.Dicussion of the survey's methodology and its key features, strengths, weaknesses.
-For instance what is the population, frame and sample.
-How is the sample recruited 
-what sampling approach is taken
-what are some of the trade offs of this 
-how is non response handled.
-(What is good and bad about it
-
-A sense of the datasets should be communicated to the reader.
-All varaiables should be throughly examined and explained.
-Explain if there were similar datasets that could have been used and why they were not.
-Why did we clean some variables?
-what the variables look like using graphs or tables discussion of the observations.
-Summary statistics should also be included
 
@@
 Thus, as we cut down the variables, they do not guarantee 100% of the vote.
-Also, because we did not exclude non-response from our dataset, for some graphs it caused difficulties in interpretation.
-There exists similar survey similar to ours, CNN exit polls, however it is not obtain the datasets they used.
+Also, we thought that excluding non-response completely would affect our understandings of data, so even though it caused difficulties in interpretation, we decided to include it except "pres16", which is our main variable whether respondents voted for Clinton and Trump.
+This is because non-response does not have meanings regarding our interests.There exists similar survey similar to ours, CNN exit polls, however it is not easy to obtain the datasets they used.
+Also, US GSS is more realistic than exit polls since people would think more rational in US GSS than exit polls.
 Now, let's get into the variables in details.
-One thing that we foucsed was whether respondents were equivalently recruited.
+One thing that we focused was whether respondents were equivalently recruited.
 To do this, we first have to look at respondents' personal information.
-If you take a look at fig(), there is not much to notice.
-Few things to notice are there are more female respondents than male, and for race, there are more White than others.
-However, following US demographics, White 76.3%, Black: 13.4%, American Indian: 1.3%, Asian: 5.9%, Hispanic: 18.5%, our sample is fairly well recruited.
+If you take a look from Figure \ref{fig:fig1} to Figure \ref{fig:fig4}.
+There exists some interesting things.
+1) number of survey participants of age in the range of 20~40 is unexpectedly low compared to other ages. \ref{fig:fig1}
+2) There are more white race than others. However, considering US race demographics, it is very reasonable. \ref{fig:fig2}
+For more details on US race demographics, https://www.census.gov/quickfacts/fact/table/US/PST045221.
+3) For gender, there is slightly more female than male, and we will see later whether this has impacts on vote or not. \ref{fig:fig3}
+4) Clinton's vote is higher than trump's vote in this survey, which contradicts the election result of 2016, reader should keep in mind that this survey is more biased towards Clinton.
+\ref{fig:fig4}
+From Figure \ref{fig:fig5} to Figure \ref{fig:fig14}, they represent respondents' thoughts on national policies as well as relation between variables.  
+Mostly they were collected from questions that whether you think that national spending on corresponding variables is about right, too much, or too little.
 
-https://www.census.gov/quickfacts/fact/table/US/PST045221.
-Considering the result above, our data sample is fairly recruited.
-
-```{r fig1, echo=FALSE}
+```{r fig1, include=FALSE, fig.cap = "Distribution of survey respondents' age group"}
 
@@
        title = "Survey Respondents' Age",
-       caption = "Figure 1: Distribution of survey respondents' age group") +
+        )+
   theme(legend.position = "none")
 ```
-
-```{r fig2, echo=FALSE}
+```{r fig2, echo=FALSE, fig.cap = "Distribution of survey respondents' race"}
 
@@
        title = "Survey Respondents' Race",
-       caption = "Figure 2: Distribution of survey respondents' race") +
+       ) +
   theme(legend.position = "none")
 ```
-
-```{r fig3, echo=FALSE}
+```{r fig3, echo = FALSE, fig.cap= "Distribution of survey respondents' gender"}
 reduced_data %>%
@@
        title = "Survey Respondents' Gender",
-       caption = "Figure 3: Distribution of survey respondents' gender") +
+       ) +
   theme(legend.position = "none")
 ```
-
-```{r fig4, echo=FALSE}
+```{r fig4, echo=FALSE, fig.cap = "Distribution of survey respondents' vote choice"}
 reduced_data %>%
@@
        title = "Surveyee's Vote Choice of 2016 Presidential Election",
-       caption = "Figure 4: Distribution of survey respondents' vote choice") +
+       ) +
   theme(legend.position = "none")
 ```
-
-```{r fig5, echo=FALSE}
+```{r fig5, echo=FALSE, fig.cap = "Opinion on national spending on improving conditions of black population"}
 reduced_data %>%
@@
        title = "Is the US Spending Enough to Improve Conditions of Black Population?",
-       caption = "Figure 5: Opinion on national spending on improving conditions of black population") +
+       ) +
   facet_wrap(~ pres16) + 
@@
 ```
-
-```{r fig6, echo=FALSE}
+```{r fig6, echo=FALSE,fig.cap = "Distribution of survey respondents' vote choice by birth place"}
 reduced_data %>%
@@
        title = "Surveyee's Vote Choice of 2016 Presidential Election by Brith Place",
-       caption = "Figure 6: Distribution of survey respondents' vote choice by birth place",
        fil = "Candidate")
 ```  
-
-```{r fig7, echo=FALSE}
+```{r fig7, echo=FALSE, fig.cap = "Distribution of survey respondents' vote by gender"}
 reduced_data %>%
@@
        y = element_blank(),
-       title = "Surveyee's Vote Choice of 2016 Presidential Election by Gender",
-       caption = "Figure 7: Distribution of survey respondents' vote by gender",)
+       title = "Surveyee's Vote Choice of 2016 Presidential Election by Gender"
+       )
 ```
-
-```{r fig8, echo=FALSE}
+```{r fig8, echo=FALSE, fig.cap = "Opinion on national spending on welfare"}
 reduced_data %>%
@@
        title = "Is the Government Spending Enough on Welfare?",
-       caption = "Figure 8: Opinion on national spending on welfare") +
+       ) +
   facet_wrap(~ pres16) + 
@@
 ```
-
-```{r fig9, echo=FALSE}
+```{r fig9, echo=FALSE, fig.cap = "Distribution of votes by degree and family income"}
 reduced_data %>%
@@
        y = element_blank(),
-       title = "Education, Income and its Effect to the Vote",
-       caption = "Figure 9: Distribution of votes by degree and family income")
+       title = "Education, Income and its Effect to the Vote"
+       )
 ```
-
-```{r fig10, echo=FALSE}
+```{r fig10, echo=FALSE, fig.cap = "Opinion on wheter the US should limit immigration to protect their culture"}
 reduced_data %>%
@@
        y = element_blank(),
-       title = "Should the Government Limit Immigration?",
-       caption = "Figure 10: Opinion on wheter the US should limit immigration to protect their culture") +
+       title = "Should the Government Limit Immigration?"
+       ) +
   facet_wrap(~pres16)
 ```
-
-```{r fig11, echo=FALSE}
+```{r fig11, echo=FALSE, fig.cap = "Surveyee's opinion on government's budget on education improvement"}
 reduced_data %>%
@@
        y = element_blank(),
-       title = "Is the Government Spending Enough on Improving Education?",
-       caption = "Figure 11: Surveyee's opinion on government's budget on education improvement")
+       title = "Is the Government Spending Enough on Improving Education?"
+       )
 ```
-
-```{r fig12, echo=FALSE}
+```{r fig12, echo=FALSE, fig.cap = "Surveyee's opinion on government's budget on arms"}
 reduced_data %>%
@@
        y = element_blank(),
-       title = "Is the Government Spending Enough on Arms and National Defense?",
-       caption = "Figure 12: Surveyee's opinion on government's budget on arms")
+       title = "Is the Government Spending Enough on Arms and National Defense?"
+       )
 ```
-
-```{r fig13, echo=FALSE}
+```{r fig13, echo=FALSE, fig.cap= "Surveyee's opinion on government's budget on assisting the poor"}
 reduced_data %>%
@@
        y = element_blank(),
-       title = "Is the Government Spending Enough to Assis the Poor?",
-       caption = "Figure 13: Surveyee's opinion on government's budget on assisting the poor")
+       title = "Is the Government Spending Enough to Assis the Poor?"
+       )
 ```
-
-```{r fig14, echo=FALSE}
+```{r fig14, echo=FALSE, fig.cap = "Surveyee's opinion on government's budget on assisting other countries"}
 reduced_data %>%
@@
        y = element_blank(),
-       title = "Is the Government Spending Enough to Assis Other Countries?",
-       caption = "Figure 14: Surveyee's opinion on government's budget on assisting other countries")
+       title = "Is the Government Spending Enough to Assis Other Countries?"
+       )
 ```
-
 ```{r, echo=FALSE}

### 2022-03-19 outputs/paper/paper.Rmd
Figure fixed and data writing

@@
 To do this, we first have to look at respondents' personal information.
-If you take a look from Figure \ref{fig:fig1} to Figure \ref{fig:fig4}.
+If you take a look from Figure \@ref(fig:fig1) to Figure \@ref(fig:fig4).
 There exists some interesting things.
-1) number of survey participants of age in the range of 20~40 is unexpectedly low compared to other ages. \ref{fig:fig1}
-2) There are more white race than others. However, considering US race demographics, it is very reasonable. \ref{fig:fig2}
+1) number of survey participants of age in the range of 20~40 is unexpectedly low compared to other ages. \@ref(fig:fig1)
+2) There are more white race than others. However, considering US race demographics, it is very reasonable. \@ref(fig:fig2)
 For more details on US race demographics, https://www.census.gov/quickfacts/fact/table/US/PST045221.
-3) For gender, there is slightly more female than male, and we will see later whether this has impacts on vote or not. \ref{fig:fig3}
+3) For gender, there is slightly more female than male, and we will see later whether this has impacts on vote or not. \@ref(fig:fig3)
 4) Clinton's vote is higher than trump's vote in this survey, which contradicts the election result of 2016, reader should keep in mind that this survey is more biased towards Clinton.
-\ref{fig:fig4}
-From Figure \ref{fig:fig5} to Figure \ref{fig:fig14}, they represent respondents' thoughts on national policies as well as relation between variables.  
+\@ref(fig:fig4)
+From Figure \@ref(fig:fig5) to Figure \@ref(fig:fig6), they represent respondents' thoughts on national policies as well as relation between variables.  
 Mostly they were collected from questions that whether you think that national spending on corresponding variables is about right, too much, or too little.
+For instance, let's take a look at \@ref(fig:fig7).
+This graph represents respondents' thoughts on U.S national spending on improving black condition, and we sorted it with our interest, Clinton or Trump.
+For more example, let's take a look at \@ref(fig:fig9).This graph represents relation between income and degree with respect to our interest.
+It generally shows that higher education you have more income you earn.
+We are going to talk detailed later in result session.
+For more information regarding variables graph please check Appndix.
 
-```{r fig1, include=FALSE, fig.cap = "Distribution of survey respondents' age group"}
+```{r fig1, echo=FALSE, fig.cap = "Distribution of survey respondents' age group"}
 

### 2022-03-19 outputs/paper/paper.Rmd
introduction

--- 2022-03-19 outputs/paper/paper.Rmd
@@
 Among numerous interesting findings from the dataset, we choose to analyze the 2016 United States presidential election.
-This slightly deviates the US GSS goal, but we thought it clearly represents current situation of US.
-However, choosing specific variables related to the election was not easy since deciding who to vote is followed by countless factors.
-After looking over the variables, we chose age,race,gender,degree,born US, Family income, national spending on education, race, arms, welfare, the poor, the other countries, social security and immigration as potential variables.
+This slightly deviates from the US GSS goal, but we thought that it can clearly represent current situations of U.S.
+However, choosing specific variables related to the election was not easy since deciding who to vote is followed by countless and complex factors.
+After looking over the variables, we chose age,race,gender,degree,born US, family income, national spending on education, race, arms, welfare, the poor, the other countries, social security and immigration as potential variables.
+
 We mainly divided the variables into two parts: personal information (Age, race, gender, degree, born US and family income) and thoughts on national policies (national spending on education, race, arms, welfare, the poor, the other countries, social security and immigration).
@@
 Thus, as we cut down the variables, they do not guarantee 100% of the vote.
-Also, we thought that excluding non-response completely would affect our understandings of data, so even though it caused difficulties in interpretation, we decided to include it except "pres16", which is our main variable whether respondents voted for Clinton and Trump.
-This is because non-response does not have meanings regarding our interests.There exists similar survey similar to ours, CNN exit polls, however it is not easy to obtain the datasets they used.
+Also, we thought that excluding non-response completely would be a bias for our understandings in data, so even though it caused difficulties in interpretation, we decided to include it except "pres16", which is our main variable whether respondents voted for Clinton and Trump.
+This is because non-response does not have meanings with respect to our interests.
+There exists similar survey similar to ours, CNN exit polls, however it is not easy to obtain the datasets they used.
 Also, US GSS is more realistic than exit polls since people would think more rational in US GSS than exit polls.
+
 Now, let's get into the variables in details.
-One thing that we focused was whether respondents were equivalently recruited.
-To do this, we first have to look at respondents' personal information.
-If you take a look from Figure \ref{fig:fig1} to Figure \ref{fig:fig4}.
-There exists some interesting things.
-1) number of survey participants of age in the range of 20~40 is unexpectedly low compared to other ages. \ref{fig:fig1}
-2) There are more white race than others. However, considering US race demographics, it is very reasonable. \ref{fig:fig2}
-For more details on US race demographics, https://www.census.gov/quickfacts/fact/table/US/PST045221.
-3) For gender, there is slightly more female than male, and we will see later whether this has impacts on vote or not. \ref{fig:fig3}
+One thing that we focused at first was whether respondents were equivalently recruited.
+To do this, we have to look at respondents' personal information.
+If you take a look from figure \@ref(fig:fig1) to \@ref(fig:fig4).
+We can see some interesting observations as followed:
+1) number of survey participants of age in the range of 20~40 is unexpectedly low compared to other ages.
+2) There are more white race than others. However, considering US race demographics, it is very reasonable. 
+For more details on US race demographics, please visit the link below: https://www.census.gov/quickfacts/fact/table/US/PST045221.
+3) For gender, there is slightly more female than male, and we will see later whether this has impacts on vote or not. 
 4) Clinton's vote is higher than trump's vote in this survey, which contradicts the election result of 2016, reader should keep in mind that this survey is more biased towards Clinton.
-\ref{fig:fig4}
-From Figure \ref{fig:fig5} to Figure \ref{fig:fig14}, they represent respondents' thoughts on national policies as well as relation between variables.  
+
+From figure \@ref(fig:fig5) to \@ref(fig:fig14), they represent respondents' thoughts on national policies as well as relation between variables.  
 Mostly they were collected from questions that whether you think that national spending on corresponding variables is about right, too much, or too little.
+For instance, let's take a look at figure \@ref(fig:fig7).
+This graph represents respondents' thoughts on U.S national spending on improving black condition, and we sorted it with our interest, Clinton or Trump.
+For more example, let's take a look at figure \@ref(fig:fig9).This graph represents relation between income and degree with respect to our interest.
+It generally shows that higher education you have more income you earn.
+We are going to talk about interesting results in more detailed later in result and discussion session.
+For more information regarding variables graph please check Appendix.
 
-```{r fig1, include=FALSE, fig.cap = "Distribution of survey respondents' age group"}
+```{r fig1,echo=FALSE, fig.cap = "Distribution of survey respondents' age group."}
 
@@
        y = element_blank(),
-       title = "Survey Respondents' Age",
-        )+
+       title = "Survey Respondents' Age") +
   theme(legend.position = "none")
 ```
-```{r fig2, echo=FALSE, fig.cap = "Distribution of survey respondents' race"}
+
+```{r fig2, fig.cap= "Distribution of survey respondents' race.", echo = FALSE}
 
@@
 ```
+
 ```{r fig3, echo = FALSE, fig.cap= "Distribution of survey respondents' gender"}
@@
 ```
-```{r fig4, echo=FALSE, fig.cap = "Distribution of survey respondents' vote choice"}
+
+```{r fig4, echo=FALSE, fig.cap = "Distribution of survey respondents"}
 reduced_data %>%
@@
 ```
+
 ```{r fig5, echo=FALSE, fig.cap = "Opinion on national spending on improving conditions of black population"}
@@
 ```
+
 ```{r fig6, echo=FALSE,fig.cap = "Distribution of survey respondents' vote choice by birth place"}
@@
 ```
+
 ```{r fig7, echo=FALSE, fig.cap = "Distribution of survey respondents' vote by gender"}
@@
 ```
+
 ```{r fig8, echo=FALSE, fig.cap = "Opinion on national spending on welfare"}
@@
 ```
+
 ```{r fig9, echo=FALSE, fig.cap = "Distribution of votes by degree and family income"}
@@
 ```
+
 ```{r fig10, echo=FALSE, fig.cap = "Opinion on wheter the US should limit immigration to protect their culture"}
@@
 ```
+
 ```{r fig11, echo=FALSE, fig.cap = "Surveyee's opinion on government's budget on education improvement"}
@@
 ```
+
 ```{r fig12, echo=FALSE, fig.cap = "Surveyee's opinion on government's budget on arms"}
@@
 ```
+
 ```{r fig13, echo=FALSE, fig.cap= "Surveyee's opinion on government's budget on assisting the poor"}
@@
 ```
+
 ```{r fig14, echo=FALSE, fig.cap = "Surveyee's opinion on government's budget on assisting other countries"}
@@
 ```
+
 ```{r, echo=FALSE}
@@
 # Appendix {-}
+Figure \@ref(fig:fig6) represents people who were born in US and their votes on the candidates. Most respondents were born in US.
+Figure \@ref(fig:fig8) represents national spending on welfare. People who advocated more funds in improvement of welfare shows preference in Clinton and people who did not shows preference in Trump. 
+Figure \@ref(fig:fig10) represents whether U.S should limit immigration in order to protect their own cultures or not.
+Even though many answered "Not Applicable", we can observe two different patters in Clinton and Trump.
+Figure \@ref(fig:fig11) represents  national spending on education with votes. People who want more funds in education improvement shows preference in Clinton and vice versa.
+Figure \@ref(fig:fig12) represents national spending on the arms with respect to votes for the candidates. People who thought too much clealry shows the preference toward Clinton. 
+Figure \@ref(fig:fig13) represents national spending on the poor with respect to respondents' votes on the candidates.
+From "About right" to "Too much", there exits more Trump supporters than Clinton's and vice versa.
+Figure \@ref(fig:fig14) represents national spending on aids to foreign countries with respect to the votes for candidates .
+There exists more Trump supporters in "Too much" comparing to Clinton's supporters.
 

### 2022-03-19 outputs/paper/paper.Rmd
edit some

@@
 # Introduction
-1) broader context to motivate
-2) some detail about hat the papaer is about
-3) a clear gap that needs to be filled
-4) what was done
-5) what was found
-6) why it is important
-7) The structure of the paper
-(3~4 paragraphs)
-Donald Trump won the 2016 United States presidential election. We are not US citizens. 
-However, considering
+Donald Trump won the 2016 United States presidential election.
+From what we remembered, it was quite surprising considering his image outside of U.S or even in U.S.
+Also, from what we remembered, before the election, Hillary Clinton was winning in most surveys as well.
+There is a famous quote that we learn from the past.
+Therefore, we thought that even though we are living in 2022, it would be meaningful to analyze the election so that we get to know what U.S citizens thoughts, or in broad sense thoughts trends in nowadays and what factors they or we considered most important to decide our representative.
+
+This paper is going to explore mainly possible variables that potentially had impacts on deciding factors in the election as well as their analysis.
+However, making a decision that this variable affects U.S citizens to lean on Trump side or Clinton side is not easy because there is nothing in this world absolute.
+Thus, as we mentioned above, we are going to focus on general flows of results instead of defining variables as absolute factors of the demarcation.
+Following that, this paper has structures: Data, Results, Discussion, Appendix and supplementary survey.
+
+We have obtained dataset from U.S GSS, and sorted variables that we consider as potential factors affecting the election.
+Throughout analysis of relation between the variables and respondents' votes on the election, we could observe that people who have conservative thoughts on equality and welfare preferred Trump, and people who do not preferred Clinton.
+This may sound obvious considering their parties, however in result session, reader can observe that what specific areas or fields they become more conservative and progressive.
+We believe that this situation is not only happening in U.S, but it also happens all over the world.
+This claim is going to be conducted with supplementary survey session.
 
@@
 We can see some interesting observations as followed:
+
 1) number of survey participants of age in the range of 20~40 is unexpectedly low compared to other ages.
+
 2) There are more white race than others. However, considering US race demographics, it is very reasonable. 
-For more details on US race demographics, please visit the link below: https://www.census.gov/quickfacts/fact/table/US/PST045221.
+For more details on US race demographics, please visit the link: https://www.census.gov/quickfacts/fact/table/US/PST045221.
+
 3) For gender, there is slightly more female than male, and we will see later whether this has impacts on vote or not.
+
 4) Clinton's vote is higher than trump's vote in this survey, which contradicts the election result of 2016, reader should keep in mind that this survey is more biased towards Clinton.
@@
 From figure \@ref(fig:fig5) to \@ref(fig:fig14), they represent respondents' thoughts on national policies as well as relation between variables.
-Mostly they were collected from questions that whether you think that national spending on corresponding variables is about right, too much, or too little.
-For instance, let's take a look at figure \@ref(fig:fig7).
+Mostly they were collected from questions that whether you think that national spending on corresponding variables is about right, too much, or too little.For instance, let's take a look at figure \@ref(fig:fig7).
 This graph represents respondents' thoughts on U.S national spending on improving black condition, and we sorted it with our interest, Clinton or Trump.
@@
 # Appendix {-}
-Figure \@ref(fig:fig6) represents people who were born in US and their votes on the candidates. Most respondents were born in US.
-Figure \@ref(fig:fig8) represents national spending on welfare. People who advocated more funds in improvement of welfare shows preference in Clinton and people who did not shows preference in Trump. 
-Figure \@ref(fig:fig10) represents whether U.S should limit immigration in order to protect their own cultures or not.
+1) Figure \@ref(fig:fig6) represents people who were born in US and their votes on the candidates. Most respondents were born in US.
+
+2) Figure \@ref(fig:fig8) represents national spending on welfare.
+People who advocated more funds in improvement of welfare shows preference in Clinton and people who did not shows preference in Trump.
+
+3) Figure \@ref(fig:fig10) represents whether U.S should limit immigration in order to protect their own cultures or not.
 Even though many answered "Not Applicable", we can observe two different patters in Clinton and Trump.
-Figure \@ref(fig:fig11) represents  national spending on education with votes. People who want more funds in education improvement shows preference in Clinton and vice versa.
-Figure \@ref(fig:fig12) represents national spending on the arms with respect to votes for the candidates. People who thought too much clealry shows the preference toward Clinton. 
-Figure \@ref(fig:fig13) represents national spending on the poor with respect to respondents' votes on the candidates.
+
+4) Figure \@ref(fig:fig11) represents  national spending on education with votes. People who want more funds in education improvement shows preference in Clinton and vice versa.
+
+5) Figure \@ref(fig:fig12) represents national spending on the arms with respect to votes for the candidates. People who thought too much clealry shows the preference toward Clinton.
+
+6) Figure \@ref(fig:fig13) represents national spending on the poor with respect to respondents' votes on the candidates.
 From "About right" to "Too much", there exits more Trump supporters than Clinton's and vice versa.
-Figure \@ref(fig:fig14) represents national spending on aids to foreign countries with respect to the votes for candidates .
+
+7) Figure \@ref(fig:fig14) represents national spending on aids to foreign countries with respect to the votes for candidates .
 There exists more Trump supporters in "Too much" comparing to Clinton's supporters.

### 2022-03-20 outputs/paper/paper.Rmd
results

@@
   drop_na(age_group) %>%
-  ggplot(aes(x = age_group, fill = age_group)) +
+  ggplot(aes(x = age_group, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
   geom_bar() +
+  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
   coord_flip() +
@@
        title = "Survey Respondents' Age") +
-  theme(legend.position = "none")
+  theme(legend.title = element_blank())
 ```
@@
        y = element_blank(),
-       title = "Is the Government Spending Enough to Assis the Poor?"
+       title = "Is the Government Spending Enough to Assist the Poor?"
        )
@@
 
-```{r, echo=FALSE}
+# Results
 
-# reduced_data %>%
-#   ggplot(aes(x = degree, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-#   geom_bar() +
-#   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-#   labs(fil = "Candidate")
+The questions asking personal information shows some obvious but also interesting results.
+Figure \@ref(fig:fig1) shows that younger generation did not participate in the survey as much as older generation.
+It can also be seen that while almost all age groups preferred Clinton over Trump, compared to younger generation, older age groups shows more interest to Trump.
+Females participants outnumbers male participants by significant amount as seen on Figure \@ref(fig:fig3), and many of female participants' vote went to Clinton, while male votes divided almost evenly(Figure \@ref(fig:fig6)).
+Figure \@ref(fig:fig8) shows some very interesting responses.
+Considering the distribution of votes, votes of white population are fairly split between Clinton and Trump, with more votes to Clinton, but all other races heavily favors Clinton.
+Some aspects that are not so obvious are ones involving education, family income and its correlation with the candidate preference.
+AS seen on figure \@ref(fig:fig9), People with higher education and higher income voted for Clinton while lower income and education votes are quite evenly split.
 
-# reduced_data %>%
-#   ggplot(aes(x = Family_income, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-#   geom_bar() +
-#   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-#   labs(fil = "Candidate") +
-#   coord_flip()
+The questions that involves asking their opinions of various subjects of government policies, figure \@ref(fig:fig8) shows that white population's general opinion is that the government's budget on welfare is not lacking, with many of them thinking government is spending too much, whereas other races opinions are evenly split with black community leaning more towards the opinion that it is lacking.
+We can also observe from the same figure that Clinton supporters generally think that the budget on welfare is lacking and Trump's supporters think budget is too much.
+Figure \@ref(fig:fig10) shows answers to question on immigration and whether the government should limit immigration to protect american culture, and it shows that out of the people who answered the question, majority of Clinton voters disagrees and majority of Trump voters agrees.
+We start to see a trend here that Trump supporters have more conservative opinions and Clinton supporters are more open to ideas on improvement in social equality or welfare.
+We see this trend continue on other figures.
+On Figure \@ref(fig:fig12), many Clinton voters think government is spending too much of national defense while very few Trump voters think the same, while many trump voters think they spend too little when very few Clinton voters have the same opinion.
+For the question on if they think the government is spending enough to assist other countries (Figure \@ref(fig:fig14)), while Clinton's supporters opinions are divided quite evenly between "about right", and "too much", majority of Trump supporters think that the government is spending too much.
+Figure \@ref(fig:fig11) and Figure \@ref(fig:fig13) shows where Clinton voters and Trump voters share the same opinion on government spending on education and to assist the poor.
+While majority of people who thinks that the government is spending too much on both occasions are Trump supporters, popular opinions from both parties is that the government is spending too little on education and to assist the poor.
 
-# reduced_data %>%
-#   ggplot(aes(x = nation_race, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-#   geom_bar() +
-#   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-#   labs(fil = "Candidate")
-
-# reduced_data %>%
-#   ggplot(aes(x = nation_welfare, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-#   geom_bar() +
-#   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-#   labs(fil = "Candidate")
-
-# reduced_data %>%
-#   ggplot(aes(x = social_security, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-#   geom_bar() +
-#   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-#   labs(fil = "Candidate")
-
-# reduced_data %>%
-#   ggplot(aes(x = race, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-#   geom_bar() +
-#   scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-#   labs(fil = "Candidate")
-
-```
-
-# Results
-Summary statistics, tables, graphs, images .... 
-Regression tables must not contain stars??
-
-Thesis: what triggered U.S citizens to vote Clinton and Trump.
-
-1. nation_race 
--> This session clearly shows the difference in approval rating between Hillary Clinton and 
-Donald Trump. Majority of people who thinks that national spending on improving conditions on
-race has to be increased advocated Hillary Clinton. On the other hand, people who did not agree 
-with them advocated Donald Trump. I assume that this is due to the image of Donald Trump.
-
-2. nation_arms
--> Interestingly this session shows opposite trend compared to nation_race. Aside from people who
-answered "About right" and "NA", the two bars, "Too little" and "Too much" shows us opposite trend.
-
-In general, similar to the indicators above, Hillary Clinton supporters were moved by improvement in social equality or welfare such as 
-"Assistance to the Poor", "nation welfare", "nation education" and "Assistance to the other countries".
-On the other hand, Donald Trump supporters were interested by more conservative ideas that US is spending too much on the sessions above.
+In general, just as we expected by the trend we saw during our observation of the data, Hillary Clinton supporters were moved by improvement in social equality or welfare such as "Assistance to the Poor", "nation welfare", "nation education" and "Assistance to the other countries".
+On the other hand, Donald Trump supporters were interested by more conservative ideas that US is spending too much on the sessions above, and shows their support to Trump's slogan "Make America Great Again".
 

### 2022-03-20 outputs/paper/paper.Rmd
results fix

@@
 
+\newpage
+
 # Results
-
 The questions asking personal information shows some obvious but also interesting results.
@@
 
-The questions that involves asking their opinions of various subjects of government policies, figure \@ref(fig:fig8) shows that white population's general opinion is that the government's budget on welfare is not lacking, with many of them thinking government is spending too much, whereas other races opinions are evenly split with black community leaning more towards the opinion that it is lacking.
+Figures 9-14 focuses on the questions that involves asking their opinions of various subjects of government policies and how supporters of different candidate , figure \@ref(fig:fig8) shows that white population's general opinion is that the government's budget on welfare is not lacking, with many of them thinking government is spending too much, whereas other races opinions are evenly split with black community leaning more towards the opinion that it is lacking.
 We can also observe from the same figure that Clinton supporters generally think that the budget on welfare is lacking and Trump's supporters think budget is too much.

### 2022-03-20 outputs/paper/paper.Rmd
... help

@@
 # Discussion
+In this session, we are going to explore important questions that can be raised from our findings.
+At this point, you may end up with question marks in meanings this analysis possess.
+Although this paper is a research of the past, this definitely has something to tell us about the U.S or even the world.
+Hopefully following questions helps you to understand our paper.
 
-what is done in this paper?
-what is something that we learn about the world
-what is another thing that we learn about the world
-what are some weaknesses of what was done?
-what is left to learn or how should we proceed in the future?
+## What was done in this paper?
 
-## First discussion point
+## What is something that we learn about the world?
+Even though the world has been trying hard for social welfare and equality, throughout this paper we have observed that there still exists disagreement within gender, race, wealth and age.
+U.S is just one example of this situation.
+If you extend your view to the world, you can observe this situation with no difficulties.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+## what is left to learn or how should we proceed in the future?
 
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
+## what are some weaknesses of what was done?
 
@@
 
-By proceeding with this survey, you understand that we will user your responses to 
-better understand () in the GSS. The survey is voluntary, and if you decide to 
-participate you can skip questions and withdraw at any time.
+By proceeding with this survey, you understand that we will use your responses to 
+better understand the 2016 U.S presidential election in the GSS. The survey is voluntary, and if you decide to participate you can skip questions and withdraw at any time.
 

### 2022-03-20 outputs/paper/paper.Rmd
survey writing

@@
 
-The purpose of this survey is ~~
+The purpose of this survey is to further investigate the 2016 U.S presidential election.
+We are interested in more and specific opinions regarding factors you consider important to decide who to vote than what US GSS provided.
+If you are not a US citizen, please pretend you are, or imagine that having very similar candidates in your home country.
 
@@
 ## Questions
-1. 
-2. 
-3.
 
@@
 
-Thank you ~~
+Thank you for participating in our survey. If you have any questions please do not hesitate to contact us.
+sangwoong.lee@maill.utoronto.ca and young.suk@mail.utoronto.ca
 
@@
 # References
+-- Smith, Tom W., Davern, Michael, Freese, Jeremy, and Morgan, Stephen L., General Social Surveys, 1972-2018 [machine-readable data file] /Principal Investigator, Smith, Tom W.; Co-Principal Investigators, Michael Davern, Jeremy Freese and Stephen L.
+Morgan; Sponsored by National Science Foundation.
+--NORC ed.-- Chicago: NORC, 2019.
+
+-- U.S. Census Bureau (2021). Population estimates. Retrieved from https://www.census.gov/quickfacts/fact/table/US/PST045221.

### 2022-03-20 outputs/paper/paper.Rmd
reference + discussion

--- 2022-03-20 outputs/paper/paper.Rmd
@@
 
+This report was created using `R` [@citeR] and its libraries `tidyverse` [@tidyverse], `dplyr` [@dplyr], were used to process and clean the data, and `ggplot` [@ggplot] `bookdown` [@bookdown] and `knitr` [@knitr] was used to format this report.
+
 # Data
@@
 3) For gender, there is slightly more female than male, and we will see later whether this has impacts on vote or not.
-
-4) Clinton's vote is higher than trump's vote in this survey, which contradicts the election result of 2016, reader should keep in mind that this survey is more biased towards Clinton.
 
@@
 
-In general, just as we expected by the trend we saw during our observation of the data, Hillary Clinton supporters were moved by improvement in social equality or welfare such as "Assistance to the Poor", "nation welfare", "nation education" and "Assistance to the other countries".
-On the other hand, Donald Trump supporters were interested by more conservative ideas that US is spending too much on the sessions above, and shows their support to Trump's slogan "Make America Great Again".
-
 # Discussion
@@
 
-## What was done in this paper?
+## Summary
+We have taken survey data to observe the trend in voters of 2016 presidential election of the United States.
+In general, just as we expected by the trend we saw during our observation of the data, Hillary Clinton supporters were moved by improvement in social equality or welfare such as "Assistance to the Poor", "nation welfare", "nation education" and "Assistance to the other countries".
+On the other hand, Donald Trump supporters were interested by more conservative ideas that US is spending too much on the sessions above, and shows their support to Trump's slogan "Make America Great Again".
 
-## What is something that we learn about the world?
-Even though the world has been trying hard for social welfare and equality, throughout this paper we have observed that there still exists disagreement within gender, race, wealth and age.
-U.S is just one example of this situation.
-If you extend your view to the world, you can observe this situation with no difficulties.
+## Disagreement Between Races, Gender, Ages, and Wealth
+Even though the United States and many other countries around the world have been trying hard to improve social welfare and equality, throughout this paper we have observed that there still exists disagreement within gender, race, wealth and age, and this had massive impact on politics and was crucial factor on people's decision on who to vote for as the president.
+Despite the popular movements voicing for equality between genders, and races and such in recent years, it seems that the opinions of people in different groups, are very divided.
 
-## what is left to learn or how should we proceed in the future?
-
-## what are some weaknesses of what was done?
+## Weaknesses and how we should proceed in the future
+The data from U.S GSS contained large amount of data and information and the distribution of survey respondents were quite accurate in terms of race.
+However, there were significantly more females than males and significantly less people of ages in the 20s, 30s and 40s.
+It is important, especially for the topic of this paper, to have the data on the younger generation and how they view government policies and how it effects their vote for the presidential election.
+To gain more knowledge in this trend of disagreement between people of different race, gender, age and wealth when it comes to the election, an additional survey could be made to answer WHY this disagreement is happening, by not just asking for their opinion but also asking about what they experience and witness.
+In the future, survey should be including questions that can allow us to better capture and understand the current trend in agreement and disagreement between different groups of people, it can allow us to study how it could effect the election in the future.
 
@@
 # References
+
+U.S. Census Bureau quickfacts: United States. (n.d.), from https://www.census.gov/quickfacts/fact/table/US/PST045221

### 2022-03-20 outputs/paper/paper.Rmd

@@
 # References
+-- Smith, Tom W., Davern, Michael, Freese, Jeremy, and Morgan, Stephen L., General Social Surveys, 1972-2018 [machine-readable data file] /Principal Investigator, Smith, Tom W.; Co-Principal Investigators, Michael Davern, Jeremy Freese and Stephen L.
+Morgan; Sponsored by National Science Foundation.
+--NORC ed.-- Chicago: NORC, 2019.
+
+-- U.S. Census Bureau (2021). Population estimates. Retrieved from https://www.census.gov/quickfacts/fact/table/US/PST045221.
 

### 2022-03-20 outputs/paper/paper.Rmd
reference

@@
 # References
--- Smith, Tom W., Davern, Michael, Freese, Jeremy, and Morgan, Stephen L., General Social Surveys, 1972-2018 [machine-readable data file] /Principal Investigator, Smith, Tom W.; Co-Principal Investigators, Michael Davern, Jeremy Freese and Stephen L.
+Smith, Tom W., Davern, Michael, Freese, Jeremy, and Morgan, Stephen L., General Social Surveys, 1972-2018 [machine-readable data file] /Principal Investigator, Smith, Tom W.; Co-Principal Investigators, Michael Davern, Jeremy Freese and Stephen L.
 Morgan; Sponsored by National Science Foundation.
@@
 
--- U.S. Census Bureau (2021). Population estimates. Retrieved from https://www.census.gov/quickfacts/fact/table/US/PST045221.
-
-U.S. Census Bureau quickfacts: United States. (n.d.), from https://www.census.gov/quickfacts/fact/table/US/PST045221
+U.S. Census Bureau (2021). Population estimates. Retrieved from https://www.census.gov/quickfacts/fact/table/US/PST045221.

### 2022-03-20 outputs/paper/paper.Rmd

@@
 From what we remembered, it was quite surprising considering his image outside of U.S or even in U.S.
-Also, from what we remembered, before the election, Hillary Clinton was winning in most surveys as well.
+Also, before the election, Hillary Clinton was ahead in most surveys as well.
 There is a famous quote that we learn from the past.
-Therefore, we thought that even though we are living in 2022, it would be meaningful to analyze the election so that we get to know what U.S citizens thoughts, or in broad sense thoughts trends in nowadays and what factors they or we considered most important to decide our representative.
+We thought that even though we are living in 2022, it would be meaningful to analyze the election so that we get to know what U.S citizens thoughts, or in broader sense, thoughts trends in nowadays and what factors we considered most important to decide our representative.
 
-This paper is going to explore mainly possible variables that potentially had impacts on deciding factors in the election as well as their analysis.
-However, making a decision that this variable affects U.S citizens to lean on Trump side or Clinton side is not easy because there is nothing in this world absolute.
+This paper is mainly going to explore possible variables that potentially had impacts on decision regarding the election as well as their analysis.
+However, making a choice for choosing which variables affected U.S citizens to lean on Trump side or Clinton side was not easy because there is nothing in this world absolute.
 Thus, as we mentioned above, we are going to focus on general flows of results instead of defining variables as absolute factors of the demarcation.
@@
 
-We have obtained dataset from U.S GSS, and sorted variables that we consider as potential factors affecting the election.
-Throughout analysis of relation between the variables and respondents' votes on the election, we could observe that people who have conservative thoughts on equality and welfare preferred Trump, and people who do not preferred Clinton.
-This may sound obvious considering their parties, however in result session, reader can observe that what specific areas or fields they become more conservative and progressive.
-We believe that this situation is not only happening in U.S, but it also happens all over the world.
+We have obtained dataset from U.S GSS, and sorted variables that we consider as potential factors affected the election.
+Throughout the analysis of relation between the variables and respondents' votes on the election, we could observe that people who had conservative thoughts on equality and welfare preferred Trump, and people who did not preferred Clinton.
+This may sound obvious considering their parties.
+In the result session, reader can observe that what specific areas or fields the respondents considered important.
+We believe that this situation is not only happening in U.S, but it is also happening all over the world.
 This claim is going to be conducted with supplementary survey session.
@@
 Usually, it is conducted in person, however because of pandemic, this time it was conducted via online.
-Among numerous interesting findings from the dataset, we choose to analyze the 2016 United States presidential election.
-This slightly deviates from the US GSS goal, but we thought that it can clearly represent current situations of U.S.
+Among numerous interesting findings from the dataset, we chose to analyze the 2016 United States presidential election.
+Although this slightly deviated from the US GSS goal, we thought that it can clearly represent current situations of U.S.
 However, choosing specific variables related to the election was not easy since deciding who to vote is followed by countless and complex factors.
@@
 As I mentioned above, deciding who to vote is followed by very complex system of thoughts.
-Thus, as we cut down the variables, they do not guarantee 100% of the vote.
-Also, we thought that excluding non-response completely would be a bias for our understandings in data, so even though it caused difficulties in interpretation, we decided to include it except "pres16", which is our main variable whether respondents voted for Clinton and Trump.
-This is because non-response does not have meanings with respect to our interests.
-There exists similar survey similar to ours, CNN exit polls, however it is not easy to obtain the datasets they used.
-Also, US GSS is more realistic than exit polls since people would think more rational in US GSS than exit polls.
+Thus, as we cut down the variables, they do not 100% guarantee of the vote.
+Also, we thought that excluding non-response completely would turn into a bias for our understandings in data, so even though it caused difficulties in interpretation, we decided to include it except "pres16", which is our main variable whether respondents voted for Clinton and Trump.
+This is because non-response does not have convey meaningful findings with respect to our interests.
+There exists a survey similar to ours, CNN exit polls, however it is not easy to obtain the datasets they used.
+Also, US GSS is more realistic than exit polls since people would think more rational with sufficient time to think in US GSS than exit polls.
 
 Now, let's get into the variables in details.
-One thing that we focused at first was whether respondents were equivalently recruited.
-To do this, we have to look at respondents' personal information.
+One thing that we focused at first was whether respondents were equivalently recruited or not.
+To do this, you have to look at respondents' personal information.
 If you take a look from figure \@ref(fig:fig1) to \@ref(fig:fig4).
-We can see some interesting observations as followed:
+You can see some interesting observations as followed:
 
@@
 From figure \@ref(fig:fig5) to \@ref(fig:fig14), they represent respondents' thoughts on national policies as well as relation between variables.
-Mostly they were collected from questions that whether you think that national spending on corresponding variables is about right, too much, or too little.For instance, let's take a look at figure \@ref(fig:fig7).
+Mostly they were collected from questions that whether you think that national spending on corresponding variables is about right, too much, or too little.For instance, let's take a look at figure \@ref(fig:fig5).
 This graph represents respondents' thoughts on U.S national spending on improving black condition, and we sorted it with our interest, Clinton or Trump.
 For more example, let's take a look at figure \@ref(fig:fig9).This graph represents relation between income and degree with respect to our interest.
-It generally shows that higher education you have more income you earn.
+It generally shows that higher education you have, more income you earn.
 We are going to talk about interesting results in more detailed later in result and discussion session.
@@
 In this session, we are going to explore important questions that can be raised from our findings.
-At this point, you may end up with question marks in meanings this analysis possess.
+At this point, you may end up with question marks in meanings this paper possess.
 Although this paper is a research of the past, this definitely has something to tell us about the U.S or even the world.
@@
 
-Our supplementary survey is available here: LINK
+Our supplementary survey is available here: https://docs.google.com/forms/d/e/1FAIpQLScN9ySPIn2W2DZ-Cs5GBpgXRcw79l1oNkqym3ARWWzIy-8uOg/viewform?usp=sf_link
 

### 2022-03-20 outputs/paper/paper.Rmd

@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "First sentence (what was done). Second sentence(what was found). Third sentence (why this matters). Fourth sentence."
+abstract: "This paper has been examined possible variables that influenced U.S citizens' choice for the 2016 U.S presidential election.
+We have found that there exists significantly different opinions regarding gender, wealth, race and age.
+This is important in a sesne that it shows thought trends as well as crucial decision indicators for voting in U.S.,With further survey at the end of this paper, we are going to explore whether this is happening only in U.S or not."
 output:
@@
 
-## Questions
-
 ## End Page

### 2022-03-20 outputs/paper/paper.Rmd
check

@@
 ---
-title: "Numerous variables relating to 2016 election in 2021 US GSS^[Code and data are available at: https://github.com/SANGWOONG-LEE/2016electionanalysis]"
+title: "Disagreement Regarding 2016 Presidential Election in the United States^[Code and data are available at: https://github.com/SANGWOONG-LEE/2016electionanalysis]"
 subtitle: "Possible factors effecting 2016 election"
@@
 However, choosing specific variables related to the election was not easy since deciding who to vote is followed by countless and complex factors.
-After looking over the variables, we chose age,race,gender,degree,born US, family income, national spending on education, race, arms, welfare, the poor, the other countries, social security and immigration as potential variables.
+After looking over the variables, we chose age, race, gender, degree, born in US, family income, national spending on education, race, arms, welfare, the poor, the other countries, social security and immigration as potential variables.
 
@@
 
-1) number of survey participants of age in the range of 20~40 is unexpectedly low compared to other ages.
+1) Number of survey participants of age in the range of 20~40 is unexpectedly low compared to other ages.
 
-2) There are more white race than others. However, considering US race demographics, it is very reasonable. 
+2) There are significantly more white population compared other races. However, considering US race demographics, it is very reasonable. 
 For more details on US race demographics, please visit the link: https://www.census.gov/quickfacts/fact/table/US/PST045221.
@@
 From figure \@ref(fig:fig5) to \@ref(fig:fig14), they represent respondents' thoughts on national policies as well as relation between variables.
-Mostly they were collected from questions that whether you think that national spending on corresponding variables is about right, too much, or too little.For instance, let's take a look at figure \@ref(fig:fig5).
+Mostly they were collected from questions that whether you think that national spending on corresponding variables is about right, too much, or too little.
+For instance, let's take a look at figure \@ref(fig:fig5).
 This graph represents respondents' thoughts on U.S national spending on improving black condition, and we sorted it with our interest, Clinton or Trump.
@@
 Figure \@ref(fig:fig1) shows that younger generation did not participate in the survey as much as older generation.
-It can also be seen that while almost all age groups preferred Clinton over Trump, compared to younger generation, older age groups shows more interest to Trump.
+It can also be seen that while almost all age groups preferred Clinton over Trump, compared to younger generation, older age groups show more interest to Trump.
 Females participants outnumbers male participants by significant amount as seen on Figure \@ref(fig:fig3), and many of female participants' vote went to Clinton, while male votes divided almost evenly(Figure \@ref(fig:fig6)).
 Figure \@ref(fig:fig8) shows some very interesting responses.
-Considering the distribution of votes, votes of white population are fairly split between Clinton and Trump, with more votes to Clinton, but all other races heavily favors Clinton.
+Considering the distribution of votes, votes of white population are fairly split between Clinton and Trump, with more votes to Clinton, but all other races heavily favor Clinton.
 Some aspects that are not so obvious are ones involving education, family income and its correlation with the candidate preference.
@@
 
-Figures 9-14 focuses on the questions that involves asking their opinions of various subjects of government policies and how supporters of different candidate , figure \@ref(fig:fig8) shows that white population's general opinion is that the government's budget on welfare is not lacking, with many of them thinking government is spending too much, whereas other races opinions are evenly split with black community leaning more towards the opinion that it is lacking.
-We can also observe from the same figure that Clinton supporters generally think that the budget on welfare is lacking and Trump's supporters think budget is too much.
-Figure \@ref(fig:fig10) shows answers to question on immigration and whether the government should limit immigration to protect american culture, and it shows that out of the people who answered the question, majority of Clinton voters disagrees and majority of Trump voters agrees.
+Figures 9-14 focuses on the questions that involves asking their opinions of various subjects of government policies and how supporters of different candidate, figure \@ref(fig:fig8) shows that white population's general opinion is that the government's budget on welfare is not lacking, with many of them thinking government is spending too much, whereas other races opinions are evenly split with black community leaning more towards the opinion that it is lacking.
+We can also observe from the same figure that Clinton supporters generally think that the budget on welfare is lacking, and Trump's supporters think budget is too much.
+Figure \@ref(fig:fig10) shows answers to question on immigration and whether the government should limit immigration to protect American culture, and it shows that out of the people who answered the question, majority of Clinton voters disagrees and majority of Trump voters agrees.
 We start to see a trend here that Trump supporters have more conservative opinions and Clinton supporters are more open to ideas on improvement in social equality or welfare.
-We see this trend continue on other figures.
+We see this trend continue on with other figures.
 On Figure \@ref(fig:fig12), many Clinton voters think government is spending too much of national defense while very few Trump voters think the same, while many trump voters think they spend too little when very few Clinton voters have the same opinion.

### 2022-03-20 outputs/paper/paper.Rmd
directory cleaning

@@
 ```{r, include=FALSE, echo=FALSE}
+# Read the reduced and cleaned data, modification done in data-preparation.r file
 reduced_data <- read_csv("outputs/data/prepared_gss.csv")
@@
 
+# Change the order of variables for graphing purposes
 reduced_data$Family_income <- factor(reduced_data$Family_income, 

### 2022-03-20 outputs/paper/paper.Rmd
edit

@@
 There is a famous quote that we learn from the past.
-We thought that even though we are living in 2022, it would be meaningful to analyze the election so that we get to know what U.S citizens thoughts, or in broader sense, thoughts trends in nowadays and what factors we considered most important to decide our representative.
+We thought that even though we are living in 2022, it would be meaningful to analyze the election so that we get to know what U.S citizens thought, or in broader sense, thoughts trends in nowadays and what factors we considered most important to decide our representative.
 
 This paper is mainly going to explore possible variables that potentially had impacts on decision regarding the election as well as their analysis.
-However, making a choice for choosing which variables affected U.S citizens to lean on Trump side or Clinton side was not easy because there is nothing in this world absolute.
+However, making a choice for  which variables affected U.S citizens to lean on Trump side or Clinton side was not easy because there is nothing in this world absolute.
 Thus, as we mentioned above, we are going to focus on general flows of results instead of defining variables as absolute factors of the demarcation.
@@
 Also, US GSS is more realistic than exit polls since people would think more rational with sufficient time to think in US GSS than exit polls.
+Ext polls can show crucial factors, but it lackes the ability of general thoughts.
 

### 2022-03-20 outputs/paper/paper.Rmd
survey

--- 2022-03-20 outputs/paper/paper.Rmd
@@
 
+```{r echo=FALSE, out.width='50%'}
+# insert pictures of survey questions
+include_graphics(paste(getwd(),"/inputs/survey/survey1.png", sep = ""))
+include_graphics(paste(getwd(),"/inputs/survey/survey2.png", sep = ""))
+include_graphics(paste(getwd(),"/inputs/survey/survey3.png", sep = ""))
+include_graphics(paste(getwd(),"/inputs/survey/survey4.png", sep = ""))
+```
+
 ## End Page
